### 2010-06-18 doc/proposition-these-matthias.tex
ma proposition de thèse
git-svn-id: http://gasp.dev.regis-gianas.com/svn/trunk@98 403e655e-980a-47bd-8c60-1ee5348bb6b7

\documentclass[twoside,a4paper,12pt]{article}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{mathpartir}
\usepackage{color}
\usepackage[pdftex,backref=page,colorlinks=true]{hyperref}
\usepackage{amsmath, amstext, amsthm, amsfonts}
\usepackage{stmaryrd}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{fancyhdr}
\input xy
\xyoption{all}
\usepackage[protrusion=true,expansion=true]{microtype}

\newcommand{\Coq}{\textsf{Coq} }
\newcommand{\Matita}{\textsf{Matita} }
\newcommand{\letin}[3]{\textsf{let}\ #1=#2\ \textsf{in}\ #3}
\newcommand{\To}{\longrightarrow}

\newtheorem{example}{Example}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}

\definecolor{bwgreen}{rgb}{0.183,1,0.5}
\definecolor{bwmagenta}{rgb}{1,0.169,0.909}
\definecolor{bwblue}{rgb}{0.317,0.161,1}
\hypersetup{
  linkcolor=blue,
  citecolor=blue
}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0.3pt}
\renewcommand{\footrulewidth}{0pt}

\fancyhf{}
\fancyhead[LE,RO]{\bfseries\thepage}
\fancyhead[LO]{\small\slshape\rightmark}
\fancyhead[RE]{\small\slshape\leftmark}
\renewcommand{\sectionmark}[1]
  {\markboth{\MakeUppercase{\thesection.\ \ #1}}
  {\MakeUppercase{\thesection.\ \ #1}}}

\date{\normalsize February 28, 2010}

\title{{\small Thesis Proposal --
    Dottorato di Ricerca in Informatica} \\[1em]
  Towards formalized mathematics repositories based on type theory}

\author{Matthias \textsc{Puech}\thanks{\texttt{\href{mailto:puech@cs.unibo.it}{puech@cs.unibo.it}}}\\[0.5em]
  \footnotesize \it Dipartimento di Scienze dell'Informazione ---
  Universit\`a di Bologna \\[-0.4em]
  \footnotesize \it Laboratoire PPS --- Université Paris 7--Denis Diderot \\[1em]
  \small \emph{dir.} Andrea \textsc{Asperti}\ \ {\it\&}\ \ Hugo
  \textsc{Herbelin}
  \\[1.3em]
}

\begin{document}

\maketitle
\thispagestyle{empty}
\tableofcontents

\section*{Introduction
  \markboth{INTRODUCTION}{INTRODUCTION}}
\pagenumbering{roman}

The fields of formal proofs, automated reasoning and verified
software, have made great advances in the last forty years, since the
seminal works of on the checkable proof language \textsf{Automath} by
\cite{de1970mathematical} and the automated theorem prover
\textsf{Nqthm} from \cite{boyer1988theorem}.  They appear today as a
mature technology, used not only academically but also in the
industry; the diversity of the approaches also witnesses the vivacity
of the field: there are plethora of systems implemented for various,
more-or-less specific uses, from the all-automated theorem prover for
a domain specific application to the general-purpose, interactive
mathematical assistant embedding a rich and expressive logic.

In particular, \emph{proof assistants}, also known as interactive
theorem provers, are tools intended to help the human to develop
formal proofs: the proof itself is constructed and checked by the
machine, but guided by the human in an interactive collaboration.
Among these, we count \textsf{Isabelle/HOL}, \textsf{PVS},
\textsf{Coq}, \textsf{Mizar}, \textsf{Matita} and much more. These
tools have recently enjoyed a great success, leading to rich formal
mathematics libraries and complex developments, see
\cite{asperti2009social} for a critical and historical account on
formal verification.

The factors of this success are in our opinion threefold: first,
advances on computational logic have lead to the development of
powerful and solid foundational languages of proofs, not only made
intuitive for the mathematician by borrowing their methods, but also
suitable to the exploitation of the computational power of modern
computers. Secondly, the wealth of implementation efforts has lead to
a well-understood architecture (\cite{asperti2007crafting}), and
efficient methods for the multitude of algorithmic tasks performed in
this context. Finally, the tight interaction between users and
developers of these systems (it is actually far from exceptional to be
both) has helped creating rich and usable interfaces, witness for
example the reinterpretation of the proof language of \textsf{Coq} by
\cite{gonthier2008small} originally intended as domain-specific but
inspiring new generations of systems.

However, a lot of the paradigms now taken for granted in the
development of formal mathematics are inherited from programming. For
example, most proof assistant are based on an LCF-like, procedural
tactic language for devising proofs, making the act of proving alike
to the one of programming and strayed from the common vernacular used
in informal mathematics. Moreover, the validation process of a formal
proof, stepwise and linear because modelled after the automatic
process of compilation, do not reflect the way a user would like to
interact with the system: not only would he like to make progress and
advance toward the solution (the only choice today), but also to
retract from previous declaration, refine and modify his idea\ldots\
Also, this rigidity does not leave room at all for team-work, where
several people interact on the same (large) document, possibly
modifying concurrently the main development in an incompatible way,
between each other or with respect to previous developments.

If, following \cite{dowek2007metamorphoses}, we believe that
machine-checkable proofs and the mechanization of proof-search made
mathematics enter its \emph{``industrial era''} after a long history
of handicraft, an era where unverified, one-man work is not acceptable
anymore due to the increased complexity of the manipulated concepts,
then we are in urge of finding techniques to adapt these new tools to
the practice of mathematics.

We believe that the simultaneous work on two aspects of this problem
would bring up new potentialities and ways of formalizing mathematics.
Increasing the inference power of the proof language is one of them:
most of today's proof languages are tightly bound to the underlying
logic, whereas informal mathematics use extensively and implicitly
notations, isomorphisms and shortcuts. Most proof assistants provide
the strict minimum to deal with them, usually in a more-or-less ad-hoc
way; increasing the capability of automating ``trivial'' parts of the
reasoning would be a first step to the design of a high-level, highly
ambiguous but still fully checkable proof language. A second step,
concerning the interactivity of the formal development process, is the
study of the impact of changes in large developments. A high-level
proof language comes with possibly high computational power required,
and one cannot afford then to recheck the whole development after each
change, as small and insignificant as they may be. Moreover, we want
to develop the interactive aspect of these tools, to be able to use
them even during the discovery phase, where change happen often and
deeply into the structure of the development. By analyzing finely the
dependency between concepts and even intra-concept, we hope to render
this interaction between the computer and the user, but also even
between several users.

All these tools are, in a nutshell, implementations of a particular
logic, along with a rich language to conduct the construction of
proofs, mechanisms of inference to relieve the user from the most
tedious and repetitive tasks, and input facilities. While the ideas
underlying the input of proofs are shared by most of these tools, a
particularly interesting subclass of them are those based on
\emph{type theory}. Type theory is a calculus having the notable
feature of being able to be seen dually as an intuitionistic logic and
as a programming language allowing to give and check rich
specifications to programs. This duality, also known as the
Curry-Howard correspondence, opened a whole new field of research,
both practical and theoretical.
\\[-0.5em]

We propose here to tackle the two issues developed above in the
context of proof assistants relying on type theory, and especially the
two developed in both our universities: \textsf{Matita} for the
University of Bologna, \textsf{Coq} for the University Paris 7. 
This proposal is therefore organized in
% three parts. The first is an
% overview of the history and main ideas underlying type theory and its
% impact on both programming languages and proof assistants. The second
two parts.
The first
is an account of the state-of-the art in terms of automation of
proofs, both on a proof-theoretic point of view and on existing
solutions implemented.  It finishes with some directions considered
for further work and works in progress in this area. 
%The last part
The second part
presents an ongoing work on \emph{semantic patches}, a preliminary
theoretical study for the management of changes in type theory that
could eventually serve as a basis for an integration into an existing
system. It finishes also with some possible directions for further
work.
We begin with a short overview of the architecture of the proof
assistants we plan to work on.

\newpage
\pagenumbering{arabic}

% \section{From construction to programming}

% When Brouwer proposed at the turn of the twenties its vision of an
% intuitionistic logic, what is in question is to reject all forms of
% formalism, proofs referring to thought experiments and nothing more:
% it is what he call the \emph{constructions}. The mathematician do not
% verify his proofs, but he must convince himself of their legitimacy,
% as he says:
% \begin{quote}
%   {\it [\ldots] in his most profound hearth}
%   (\cite{brouwer1927intuitionistic}, in \cite{van1967frege}).
% \end{quote}

% This profoundly subjectivist view of logics contrasts interestingly
% with the contemporary work on formalizing set theory by Zermelo, and
% Hilbert's dream of making mathematics so formal that an automatic
% process could generate all valid and interesting theorems of
% mathematics. He actually leaves almost entire the question, yet
% crucial, of the nature of these constructions. The interpretation by
% Heyting in the thirties of his work and the partial formalization of
% the notion gave birth to what we now call ``intuitionistic logic'', or
% constructive mathematics.

% Fifty years later, computer systems emerge that can verify the
% validity of proofs, totally automatically, and even efficient
% processes to discover proofs. A large amount of these tools, as we
% will see, are profoundly linked with the ideas that have emerged from
% the notion of constructive mathematics: in fact, many of them
% implement at their very core an ``intuitionistic logic'': the theory
% of types. Moreover, underlying the proofs that one can carry in these
% logics, there are real programs, that can be executed by the machine.
% How has the notion of construction shifted so much as to change nature
% almost entirely? It is the question we will try to explore here.

% For this, we will draw the material for this analysis in two, strongly
% dependent discoveries: one is the Curry-Howard correspondence which
% appears today as the first tangible incarnation of the idea of
% construction, the second is one of its sequel, that gave birth to a
% whole field of research at the boundaries between Mathematics, Logics
% and Computer Sciences: type theory.

% Finally, we will expose the formal definitions, the variants and some
% of the issues regarding the incarnation of type theory that will be
% the framework of this proposal: the Calculus of Constructions. It is
% this type theory that is implemented in the proof assistants
% \textsf{Coq} and \textsf{Matita}.

% \subsection{Constructive mathematics}

% We shall try in this section to give some historical references and
% clues that will help us comprehend this paradoxical development of
% history: how, from the subjectivism of Brouwer, could emerge these
% tools, that witness such an externalist vision of Logic?  How this
% intimate, almost mystical experience -- the construction -- could be
% subject to such a change of nature as to become a process, an
% algorithm for the totally external verification of proofs?

% \subsubsection{The constructivist program}

% At the turn of the 20\textsuperscript{th} century, the dominant
% ideology in mathematical logic was that of \emph{formalism}. This
% theory, as advocated among others by the mathematicians Hilbert,
% Bernays and von Neumann, holds that mathematics, and in particular
% logic, are nothing but a game on formulas and statements, and that
% theorems and their proofs can be assimilated to a bare manipulation of
% strings. For example, Euclidean geometry can be thought as a game,
% whose goal is to transform a series of axioms by a number of
% transformations into a given statement, and once this goal is
% attained, the statement has been proven. It is then a theory about,
% and only about axiomatic systems: for these mathematicians, the system
% of axioms and rules considered is of practically no importance except
% for its intrinsic properties, and the meaning of the actual choice of
% a system is (almost) totally irrelevant.

% In spite of the many deceptions that arose at the beginning of the
% century when attempting to devise a general enough formalism -- among
% others, Frege's \emph{Begriffschrift} attempt to build a universal
% formal language was doomed by Russell's paradox, see
% \cite{van1967frege} -- this view was so much anchored in its time that
% David Hilbert, maybe the earliest proponent of formalism, proposed his
% \emph{program}. From the ``foundational crisis'' mathematics were
% suffering, that is the search for a proper foundation for mathematics,
% he proposed to ground all existing theories to a finite, complete set
% of axioms, and provide a proof that these axioms were consistent, in
% order to provide mathematics with a universal, solid foundation: a
% complete and consistent axiomatization of \emph{all of mathematics}.
% However, a few years later, this program was mathematically shown to
% be unattainable by Gödel's incompleteness theorem.
% \\[-.5em]

% In this context, \emph{constructivism} emerged as a strong opponent to
% the formalist school. Lead by Brouwer, Markov, Bishop and late Hilbert
% and Bernays with their ``finitism'', constructivism, as a
% philosophical view on the existence of the mathematical objects,
% asserts that it is necessary to find (or "construct") a mathematical
% object in order to prove that it exists.  Therefore, the ``game''
% of applying inference rules from axioms is not necessarily a proof of
% existence of an object: only the \emph{construction} -- the exhibit of
% such an object -- is able to convince the mathematician of its
% validity. Thus, the formalist notion of inference disappears almost
% entirely here: the choice of the system is not only the only valuable
% question, since it would make no sense to conduct a proof in a system
% that doesn't allow to perform the construction of a witness of
% existence of the object, but it also looses all reason to be: only one
% system is capable of producing such a witness to perform constructive
% mathematics.

% In particular, one form of reasoning is therefore strongly rejected:
% when one assumes that an object does not exist and derives a
% contradiction fromc that assumption, one still has not found the object
% and therefore not proved its existence, according to constructivists:
% the \emph{law of excluded middle} is therefore rejected. Often,
% constructivism is assimilated with this rejection; it is in fact only
% a consequence of the foundation of the notion of proofs on the
% construction.

% Brouwer's \emph{intuitionism} is one of the form taken by the
% constructivists' thesis statements. It was strongly influenced by
% Kant's notion of intuition and redefined the notion of truth as it was
% considered by traditional mathematics. In classical logic like in set
% theory, truth is a static property of a proposition (or as Brouwer
% prefers to name it after Kant, a \emph{judgement}): it is either true
% or false, depending on its \emph{real nature}. This is expressed by
% the interpretation of the logical statements in a model of truth
% values: traditionally, the meaning of a proposition was interpreted in
% an algebra of two values, true and false. On the contrary, the major
% methodological turning point of Brouwer is to abandon the semantical
% interpretation of propositions. He suggests that this usual denotation
% of propositions as truth values is simply inadequate, this for four
% reasons:
% \begin{itemize}
% \item It is first a philosophical platitude: Tarski for example
%   interprets the conjunction as: $A\wedge B$ is true if $A$ is true
%   and $B$ is true.  The \emph{meta} operator ``and'' doesn't help us
%   to understand $\wedge$ in any way.
% \item Justify the quantification in this way is undecidable in most
%   cases, since we have to verify it on the whole ``domain of the
%   discourse''.
% \item Fermat's theorem denotes in this case the same object as
%   $2+2=4$. It is really reasonable to let so little room to the notion
%   of proof?
% \item Finally, this interpretation simply doesn't scale to
%   higher-order logic.
% \end{itemize}

% He takes a much more subjectivist approach to the notion of proof. For
% him, it is even close to the juridical notion of \emph{evidence}: a
% proof of $A$ is the set of processes that make one be convinced of
% $A$. It is in some way an incriminating evidence, an object the
% observer -- the judge or the mathematician -- can convince himself,
% either directly (the immediate proof), or through a series of steps, a
% mechanism (the mediate proof, like the mediate jurisdiction) that can
% be ``replayed'' at will, as many times as one needs to convince
% himself. The intuitionistic slogan of Brouwer's ``there are no
% unexperienced truth'' takes then all its meaning. The experience, that
% Brouwer call construction, is the only evidence that constitutes
% truth. In this sense, the act of ``proving'' that $12$ is not a prime
% number is to exhibit the computation of $4*3$ as an evidence.

% Heyting, and later Kolmogorov will formulate this idea more precisely
% (in some of the articles in \cite{benacerraf1983philosophy}): the
% proposition (statement, judgement) must be understood as a problem, an
% expectation or an intention. Better, a task. To prove a statement (a
% problem, a task) is to know what constitutes a solution to its
% problem, a fulfilment to its task. Heyting and Kolmogorov
% (\cite{heyting1971intuitionism}) even propose a full interpretation of
% the logical statements in terms of these analogy: the famous BHK
% interpretation: \newcommand{\subst}[2]{[{#1}/{#2}]}
% \begin{align*}
%   \llbracket A\Rightarrow B \rrbracket &= f \in \llbracket
%   A\rrbracket\to \llbracket B\rrbracket\\
%   \llbracket\forall x\in A,\ B\rrbracket &= f \in
%   \{\llbracket B_x\rrbracket\}_{x\in \llbracket A\rrbracket} \\
%   \llbracket \exists x\in A,\ B\rrbracket &= \langle a,b\rangle \in
%   \llbracket A\rrbracket\times \llbracket B\subst{x}{a}\rrbracket \\
%   \llbracket A\wedge B\rrbracket &= \langle a, b \rangle \in
%   \llbracket A\rrbracket\times \llbracket B\rrbracket
% \end{align*}

% This interpretation gave birth to the first formalization by Heyting
% of the intuitionistic logic. It is nowadays presented as a natural
% deduction, minus the rule \emph{Reduction ad absurdum} which justifies
% the excluded middle.

% This interpretation put one idea forward that takes the form of two
% properties. A logic that would satisfy this interpretation will see
% its connectives $\exists$ and $\vee$ act more precisely, with more
% control than in usual, classical logics: they are the properties of
% the witness and of the disjunction.
% \begin{property}[Witness]
%   From a proof of a statement of the form $\exists x, P(x)$, one can
%   \emph{extract} a \emph{witness} $a$ such that $P(a)$ is provable.
% \end{property}
% \begin{property}[Disjunction]
%   From a proof of a statement of the form $A\vee B$, one can decide
%   which of $A$ or $B$ is provable.
% \end{property}

% As we can see, the excluded middle $A\vee¬A$ is obviously not provable
% in such a logic: if it were, one of $A$ or $¬A$ would be provable
% (without any other knowledge about $A$), which is obviously false.
% \\[-.5em]

% The first real incarnation of what we could call a guideline for the
% elaboration of logics, was discovered in the sixties, almost by
% accident, by linking two notions that had previously no commonalities,
% the $\lambda$-calculus and the intuitionistic logic. This link
% actually turned out to be an exact and astonishing correspondence.

% \subsubsection{The Curry-Howard correspondence}

% In the 1930s, Church introduced, as part of an investigation into the
% foundation of mathematics, a very simple yet powerful formal system to
% reason on function definition and their application: the
% $\lambda$-calculus.

% After the failure of Frege's (variant of) na\"ive set theory,
% logicians were left with the difficult problem of finding a natural
% yet consistent formulation of set theory. Two direction were taken
% from this point. Zermelo and Fr\"ankel on one side modified Frege's
% formulation to obtain ZF, an axiomatic set theory. Church on the other
% side, took the approach of expressing higher-order logic (a set
% theory) with the help of a minimal language for functions. In his
% seminal article, \cite{church1940formulation} introduced, with the
% help of the $\lambda$-calculus, a formulation of classical,
% higher-order logic based on first-order logic with the addition of
% only three symbols: $\lambda$ which represents an anonymous function,
% the application of a function to its argument, and the arrow $\to$ to
% express what he call the \emph{type} of these functions. This
% formulation is called the Simple Theory of Types. It has the
% peculiarity of introducing, for the first time explicitly, a notion of
% computation or \emph{reduction}: the $\beta$-reduction.

% This reduction is at the heart of the higher-order character of his
% logics: in fact, it mimics the behavior of an axiom of set theory, the
% \emph{comprehension scheme}: \[ \forall x_1\ldots x_n, \exists y,
% \forall z (z\in y \Leftrightarrow P)\] This scheme states that it is
% possible to define sets \emph{in comprehension}, i.e. by a property on
% its elements. At each property corresponds a set. In fact, by
% skolemizing this scheme and introducing the notation $\{ x\ |\ P \}$
% for the function associated to $y$, we obtain: \[ \forall x_1\ldots
% x_n, z, (z \in \{z\ |\ P\} \Leftrightarrow P)\] This is read: if I
% constructed the set $z$ from the property $P$, and if in fact, $z$ is
% in this set, then $P$ is true.

% Some years later, Church, extracting the computational content of his
% logic, was left with a very simple formal system that had nothing to
% do with logic anymore, but with computational complexity:
% \[ t\ ::=\ x\ |\ \lambda x. t\ |\ t\ t \qquad\qquad (\lambda x. t)\ u
% \to t[x/u]\]. This systems supposedly allowed to represent \emph{any}
% kind of computation, and still today is a touchstone for the notion of
% computation. One problem with it is that, viewed as a set theory, it
% allows to define the set of all sets that do not belong to themselves,
% Russel's paradox: \[ \{x\ |\ ¬x\in x\}\in\{x\ |\ ¬x\in x\}\]. Thus,
% the above system is not only logically inconsistent but Russel's
% paradox is also a counter-example to its termination: the
% corresponding term $(\lambda x. x\ x)(\lambda x. x\ x)$ reduces on
% itself \emph{ad infinitum}. Obviously, the problem with the above
% system is that it allows the variable $x$ is applied to itself, which
% lead Church to put each variable in a distinct \emph{type}. A variable
% denoting a function would have a type $A\to B$, and could be applied
% only to an argument of type $A$. This lead to the Simple Type Theory,
% an isolation of a subset of well-behaved terms, the typed terms. They
% are defined inductively as follows:
% \begin{mathpar}
%   \infer[Init]{(x:A)\in\Gamma}{\Gamma\vdash x:A} \and
%   \infer[Abs]{\Gamma, x:A\vdash t:B}
%   {\Gamma\vdash\lambda x. t : A\to B} \and 
%   \infer[App]{\Gamma\vdash t:A\to B \and \Gamma\vdash
%   u:A}{\Gamma\vdash t\ u: B}
% \end{mathpar}
% These three rules formed so to say the core of the functional aspect
% of the logic discovered by Church, but also a universal programming
% language, in which all possible programs are expressible as terms. For
% a full history of $\lambda$-calculus and their applications, see
% \cite{cardone2006history}.
% \\[-.5em]

% This idea of a well-behaved fragment of the $\lambda$-terms stayed
% contained to the study of computable functions and higher-order
% logics, for example by \cite{russell1908mathematical} in his Ramified
% Theory of Types. In the 1960's however, H. Curry and W. A. Howard
% noted a striking similarity between these three rules and those of the
% minimal fragment of intuitionistic logic: in fact, if one removes all
% information concerning terms in these rules, he obtains exactly the
% rules of minimal intuitionistic logic:
% \begin{mathpar}
%   \infer[Init]{A\in\Gamma}{\Gamma\vdash A} \and
%   \infer[Intro]{\Gamma,A\vdash B}{\Gamma\vdash A\Rightarrow B} \and
%   \infer[Elim]{\Gamma\vdash A\Rightarrow B \and \Gamma\vdash A}{\Gamma\vdash B} \and
% \end{mathpar}

% This remark, yet so simple, circulated on an unpublished note that was
% published only twenty years after by \cite{howard1980formulae}, and
% was called the \emph{proposition-as-types} paradigm. It is easily
% expanded to full first-order intuitionistic propositional logic by
% adding new connectives in the types and new constructions in the term:
% pairs and conjunctions, pattern-matching and disjunction, empty type
% and $\bot$, unit type and $\top$\ldots\ and allows then to build a
% very precise correspondence between notions of proof-theory and their
% now-equivalent in computation theory and programming language
% theory:
% \\

% \begin{center}
%   \begin{tabular}{|l|l|}
%     \hline 
%     \bf proof-theory & \bf computation \\ \hline
%     proposition & type \\ \hline
%     proof & program \\ \hline
%     deduction rules & typing rules \\ \hline
%     cut-elimination & evaluation \\ \hline
%     cut-free proof & value \\ \hline
%   \end{tabular}
% \end{center}
% \vspace{1em}
% However, the problem of extending this correspondence to full
% first-order logic, and besides to higher-order logic stayed
% entire. One of the (many) answer to this problem appeared at the end
% of the seventies under the name of \emph{Intuitionistic Type Theory}.

% \subsection{Type theory as a foundational language}

% Among the formalisms proposed to extend the Curry-Howard
% correspondence, two of them were particularly successful. One is the
% system F, discovered independently by \cite{girard1972interpr} and
% \cite{reynolds1983types}. It is a functional programming language with
% polymorphism that can be viewed logically through the Curry-Howard
% correspondence as the minimal second-order fragment of intuitionistic
% logic. The second, which will be our interest in this section is
% Martin-L\"of's Intuitionistic Type Theory
% (\cite{martin1984intuitionistic}). It differs from the previous
% formalisms in an important way, as it is \emph{dependent}, that is, it
% makes no distinction anymore between types and terms and allows
% computation to be performed not only on terms but also on types.

% We will expose this formalism and try to hint its correspondence to
% the intuitionistic program of Brouwer, but before that we will have to
% expose one of its major underlying philosophical idea, taking its
% ground directly in the notion of judgement of Kant's. We will then see
% that it allows a very general presentation, unifying most of the
% variants proposed, including system F: pure type systems.

% \subsubsection{The judgemental method}
% \label{sec:jmeth}

% To justify the notion of dependency as introduced in his type theory,
% Martin-L\"of proposes a new method for constructing a logic, that we
% will call the \emph{judgemental method} following Pfenning. It is a
% shift of focus from the notion of formula to the one of judgement,
% allowing a new flexibility in the definition and justification of the
% meaning of the rules. This method is exposed in the course notes
% \cite{martin1996meanings}.

% What is the object of a rule of inference? What are $A$ and $A\vee B$
% when I write the rule: $$\infer{A}{A\vee B}$$ The traditional answer
% to that question is that they are \emph{formulas}, or
% \emph{propositions}. They constitute in the traditional course on
% logic, e.g. when natural deduction is explained, the central notion,
% the one on which all other are depending on: inference, axioms,
% models\ldots\ A formula is an inductively defined expression,
% that is fixed before starting to define the rules of inference, and is
% generally represented as a BNF grammar:
% $$ A,B\ ::=\ P(t_1,\ldots,t_n)\ |\ A\wedge B\ |\ A\vee B\ |\ \forall x, A\ |\
% \exists x, A\ |\ \top\ |\ \bot$$ \emph{Expressions} on the contrary
% form so to say an open concept: they are tree-like objects that one
% can represent, \emph{``can recognize in its many occurrences and can
%   reproduce in many copies''} (\cite{martin1996meanings}). From this
% traditional view, we can draw an implicit order of precedence of
% concepts:
% \begin{quotation}
%   expression $<$ formula $<$ logical rule
% \end{quotation}

% This vision has a severe conceptual flaw however. When I define
% natural deduction, the \emph{signature}, i.e. the set of predicates,
% function symbols and axioms, are to be defined before the formulas
% since formulas depend on them, and \emph{a fortiori} before logical
% rules. Then, if I define the reasoning on arithmetics (natural
% deduction on Peano's arithmetic), and then set theory (natural
% deduction on ZF axioms), I define to totally disjoints theories that
% don't even share their logical connectives: we will have
% $\vee_{\text{Peano}}$ and $\vee_{\rm ZF}$ etc. And yet, these
% connectives behave exactly the same way on the formulas: they are so
% to say two unnecessarily specific versions of the same logical
% concept.

% This remark suggests a more liberal, a looser handling of the notion
% of formula. In the above conception, an important point is missing:
% when a rule is stated, it is implicitly stated that if the premises
% are true, then the conclusion must be true as well. If we make this
% character explicit, the rule above must be read as:
% \[ \infer{A\vee B\text{ is true}}{A\text{ is true}}\] The previously
% implicit modality applied to the formulas -- the fact that they are
% true, make up what we will call a \emph{judgement}. As it was done by
% Bolzano and Frege, we can even declare different forms of judgement,
% like the fact that a statement is true or false, the fact that an
% expression is a natural number\ldots\ provided that we provide a way
% to verify these forms of judgments: provided that we know what counts
% as a proof of these judgements. Introducing this intermediate concept
% of judgment allows us to liberate ourself completely from the too
% restrictive notion of formula. Let's even replace it completely with
% the open notion of expression, and constrain \emph{by the rules} their
% character of formula or proposition. We introduce two different forms
% of judgements:
% \begin{itemize}
% \item $A$ is a proposition ($A\text{ prop}$)
% \item $A$ is true ($A\text{ true}$)
% \end{itemize}
% The previous rule becomes:
% \[ \infer{A\text{ prop}\and B\text{ prop}\and A\text{ true}}{A\vee
%   B\text{ true}}\]
% So to say, the charge of work let previously to the definition of
% formula -- discriminate against well- or ill-formed propositions -- is
% deferred on the inference rules, and a derivation in the obtained
% system not only constitutes the establishment of the truthfulness of
% a formula, but also of this well-forming. For each symbol, we now have
% to give the adequate rules:
% \begin{mathpar}
%   \infer{A\text{ prop}\and B\text{ prop}}{A\vee B\text{ prop}} \and
%   \infer{A\text{ prop}}{¬A\text{ prop}} \and \ldots
% \end{mathpar}

% The syntax of formulas, deferred this way to the logical rules, allows
% a tight control of the system, because judgement types can be mixed
% together. This remark, as innocent as it seems, is in fact a central
% point of the theory of types as we will see.

% \subsubsection{Martin-L\"of type theory}

% The original presentation of Martin-L\"of's type theory
% (\cite{martin1984intuitionistic}) is an extensible theory composed of
% a core \emph{calculus of judgements}, on top of which one declares the
% logical connectives and their rules.

% It relies on two primitive notions: the first is called indifferently
% \emph{set} or \emph{type} and corresponds intuitively to the --
% intuitionistic -- idea that a type is nothing but the set of its
% values. Beware though: ``set'' is not to be understood in the broad
% sense of set theory only as ``what we can define inductively''. We
% will write following the original presentation $\in$ the membership of
% such a set. The second notion is the one of \emph{judgmental
%   equality}. It is a form of judgement that defines the set of purely
% computation substitution operations allowed on respectively the sets
% and their inhabitants. Think for the moment of this equality as a
% quotient on the other forms of judgment.

% \paragraph{Forms of judgment}

% \newcommand\set{\text{ set}}

% We introduce the following forms of judgment:
% \begin{enumerate}
% \item $A$ is a set ($A\set$)
% \item $A$ and $B$ are equal judgments ($A=B$)
% \item $a$ is an element of $A$ ($a\in A$)
% \item $a$ and $b$ are two equal elements of $A$ ($a=b\in A$)
% \end{enumerate}

% They come with constraints on the rules they will apply to: if we want
% to define a particular set $S$, $S\set$ can be introduced as a rule
% only if we introduce at the same time all rules that describe all
% elements of this set, and how to compare two elements of this set. For
% example, this is a valid introduction of a set:
% \begin{mathpar}
%   \infer{ }{\textsf{0}\in\mathbb N} \and
%   \infer{n\in\mathbb N}{\textsf{S}\ n\in\mathbb N} \and
%   \infer{ }{\textsf{0}=\textsf{0}\in \mathbb N} \and
%   \infer{n=m\in\mathbb N}{\textsf{S}\ n = \textsf{S}\ m \in\mathbb N}
% \end{mathpar}

% All of these judgements form in fact families of judgement by what is
% introduced under the name of \emph{hypothetical judgement}. An
% hypothetical judgement is a family of judgements indexed by another
% judgement, and will be written $$ (J_1)\ldots(J_n)\ J $$

% Hypothetical judgements are judgements of their conclusion $J$ whose
% proof can be made complete by the substitution of their ``holes'' by
% judgements of their hypothesis $J_1\ldots J_n$. An important novelty
% is that it is an inductive construct: an hypothesis $J_i$ is also
% supposed to be an hypothetical judgement, parameterized by all $J_j$
% with $j<i$, and this is reflected even in its statement. This
% mechanism allows a very natural expression of mathematical statements.
% \begin{example} The mathematical(ly void) statement:
%   \begin{quote}
%     Let $R$ be a relation on a set $A$, and $R*$ its reflexive
%     closure. Then for all $x\in A$, $R*(x,x)$.
%   \end{quote}
%   is translated into the hypothetical judgement:
%   \begin{quote}
%     ($A$ is a set) \\
%     ($R$ is a relation on $A$) \\
%     ($R*$ the reflexive closure of $R$) \\
%     ($x$ an element of $A$) \\
%     $R*(x,x)$
%   \end{quote}

% \end{example}

% \paragraph{Interpretation}

% You won't be surprised to learn that each of these forms of judgement
% accept several reading, as based on the BHK interpretation:

% \begin{center}
%   \begin{tabular}{l|l}
%     $\bf A\ \rm\bf set$ & $\bf a\in A$ \\ \hline
%     $A$ is a set & $a$ is an element of $A$ 
%     \emph{or} $A$ is non-empty \\\hline
%     $A$ is a proposition & $a$ is a proof of $A$ 
%     \emph{or} $A$ is true \\ \hline
%     $A$ is a problem & $a$ is a method to solve $A$
%     \emph{or} $A$ is solvable \\ \hline
%     $A$ is a specification & $a$ is a program realizing $A$
%     \emph{or} $A$ is inhabited
%   \end{tabular}
% \end{center}

% In other words, $A\set$ can be viewed (interpreted) as the judgment
% $A\text{ prop}$ introduced in \ref{sec:jmeth}. The same way, by
% ``degrading'' the judgement $a\in A$ and forgetting the $a$, it is
% read as $A\text{ true}$. These remarks have for Martin-L\"of only an
% indicative value though: they only allow us to give intuitions on
% their meaning, and do not constitute an formal interpretation: they
% will allow, once the logic fully defined, to verify that the logical
% constants introduced have the meaning we intended to give them.

% \paragraph{Typing rules}

% We can now begin to construct the system of rules for the core
% theory. All rules can be justified by the underlying construction, but
% we detail only some of them.
% \\[-.5em]

% \underline{\bf Equality rules}
% \begin{mathpar}
%   \infer[Refl/EqS]{A\set}{A=A} \and
%   \infer[Refl/EqE]{a\in A}{a=a\in A} \\
%   \infer[Sym/EqS]{A=B}{B=A} \and
%   \infer[Sym/EqE]{a=b\in A}{b=a\in A} \\
%   \infer[Trans/EqS]{A=B \\ B=C}{A=C} \and
%   \infer[Trans/EqE]{a=b\in A \\ b=c\in A}{a=c\in A} \and
% \end{mathpar}

% These rules reflect the fact that the equality is an equivalence
% relation, on elements and on sets. They are justified by the fact that
% they define a congruence on elements and sets. Remark the use of other
% kind of judgments as premises of the reflexivity rules. They ensure
% that equality is always done in a ``typed setting'': one cannot prove
% $2=2$, but only $2=2\in\mathbb N$.
% \\[-.5em]

% \label{sec:mltt}
% \underline{\bf Set conversion}
% \begin{mathpar}
%   \infer[Conv/E]{a\in A \\ A=B}{a\in B} \and
%   \infer[Conv/EqE]{a=b\in A\\ A=B}{a=b\in B}
% \end{mathpar}

% These rules state that equal elements of different sets can be
% assimilated provided that the sets are equal. They will be of crucial
% importance and are the key to the comprehension of type theory and its
% sequel, because they break the usual discrepancy made between types as
% static values and terms as dynamic entities that can be evaluated: in
% type theory, sets are also subject to computation. For example, the
% set of vectors of length $x+y$ is the same as the set of vectors of
% length $y+x$ up to conversion.
% \\

% \newcommand{\hypo}[2]{(#1)\ #2}

% \underline{\bf Substitution}
% \begin{mathpar}
%   \infer[Subst/E/S]{a\in A \\ \hypo{x\in A}{B\set}}{B\subst{x}{a}\set} \and
%   \infer[Subst/E/EqS]{a\in A \\ \hypo{x\in
%       A}{B=D}}{B\subst{x}{a}=D\subst{x}{a}} \and
%   \infer[Subst/E/E]{a\in A \\ \hypo{x\in A}{b\in B}}{b\subst{x}{a}\in
%     B\subst{x}{a}} \and
%   \infer[Subst/E/EqE]{a\in A \\ \hypo{x\in A}{b=d\in
%       B}}{b\subst{x}{a}=d\subst{x}{a}\in B\subst{x}{a}} \and
%   \infer[Subst/EqE/S]{a=c\in A \\ \hypo{x\in
%       A}{}B\set}{B\subst{x}{a}=B\subst{x}{c}} \and
% \end{mathpar}

% Substitution is the basic operation on hypothetical judgements. They
% state that an hypothetical judgement can be completed by the knowledge
% of its hypothesis; there are theoretically one rule by pair of
% judgement, but we keep only the useful ones here. Here we have the
% witness of the \emph{operationality} of this language: this mechanism
% of completion of partial knowledge forms already an algorithm for the
% reduction of a proof, or the execution of the solution to a problem in
% Martin-L\"of's words. This basic system, forms the core of
% Martin-L\"of's type theory, in which one can define the usual logical
% as well as general connectives such as the dependent products and
% sum.

% This language, by its operationality and its ability to represent
% logical concepts in an intuitionistic way, is the basis of many modern
% proof assistants. In fact, Martin-L\"of goes further and declares:

% \begin{quote} \textit{ ``If programming is understood not as the
%     writing of instructions for this or that computing machine but as
%     the design of methods of computation that it is the computer's
%     duty to execute (a difference that Dijkstra has referred to as the
%     difference between computer science and computing science), then
%     it no longer seems possible to distinguish the discipline of
%     programming from constructive mathematics. This explains why the
%     intuitionistic theory of types [\ldots], which was originally
%     developed as a symbolism for the precise codification of
%     constructive mathematics, may equally well be viewed as a
%     programming language.''} --- \cite{martin1984constructive}
% \end{quote}

% \subsection{Architecture of a proof assistant}
\subsection*{Architecture of a proof assistant}

Both our proof assistants of interest, \textsf{Coq} and
\textsf{Matita} are based on the \emph{Calculus of Inductive
  Constructions (CIC)} of \cite{paulin1996definitions}. It is a very
expressive logic, built on a variant of Martin-L\"of's type theory
with polymorphism \`a la System F and a hierarchy of predicative
universes. It supports, as in Martin-L\"of presentation of his type
theory, the inductive definitions of objects, and supports the
proofs-as-types paradigm. Among other features, it provides recursive
function definitions and an impredicative universe for dealing with
logical propositions.

Numerous systems based on a variant of this framework have been
implemented: \textsf{Agda}, \textsf{NuPRL}, \textsf{Epigram} and much
more. As time went by, the architecture of these tools has become
better and better understood. One particularly important criterion in
their conception is the so-called \emph{De Bruijn criterion}. They all
share a common layered software architecture, as sketched for
\textsf{Matita} in \cite{asperti2007crafting} for example, and are
often qualified of ``skeptical proof assistants''. We give a brief
overview of each layer from the innermost to the outermost one, as it
is implemented in \textsf{Matita} (this may vary for other systems):

\begin{description}
\item[Kernel] Their principle is to isolate a little portion of
  human-understandable code, the \emph{kernel}, only dedicated to the
  task of proof-checking, i.e. the verification of the well-typing.
  This component is responsible for the validation of all generated
  proofs, and is this way the only piece of code that must be trusted
  in order to trust the whole system.

\item[Refiner] Writing directly proofs in the language of the CIC is
  an almost impossible operation: not only would it be very hard to
  follow, but also very boring, because the language is very
  redundant. On top of the kernel, the \emph{refiner} is in charge of
  completing all logical information omitted by the user, with the
  help of dedicated mechanisms: \emph{type inference} allows you to
  omit some type constraints if they can be inferred by the system
  from the context. \emph{Coercions} simulate a common practice in the
  mathematical discourse, i.e. the use of a super-structure in place
  of one of its components (example, a group instead of its carrier
  set). The more recent \emph{canonical structures} and \emph{type
    classes} allow to avoid mentioning the context by choosing, on one
  side a canonical representative of the context (a particular
  mathematical structure) or relying on inheritance between
  structures.

\item[Library management] All proofs and definitions in these proof
  assistants are devised thanks to a mechanism of definitional
  equality (in the sense of definition expansion). All the definitions
  are stored in a library, with the ability to import and reuse
  previous development. This layer is in charge of the management of
  such a library: loading, saving, managing namespaces as well as
  modifying previous definitions.

\item[Proof management] Some of the systems mentioned allow the
  stepwise construction of proofs thanks to a library of tactics,
  which are tools to build proofs incrementally by refining the
  previous state of the proof. This layer manages the call to tactics
  in this process, and the verification of their result by the refiner
  (which in turns calls the kernel).

\item[Tactics] This layer makes up the set of specific tactics of
  proof refinement. It is usually built in a hierarchical manner, some
  tactics calling some others for specific subtasks.

\item[Disambiguation] Very often in the context of informal
  mathematical vernacular, ambiguity of the notations used happens.
  This layer is responsible for mapping potentially ambiguous
  notations and names to their logical, non-ambiguous counterpart by
  inferring it from the context.

\item[User interface] Then, some logic-independent layers are in
  charge of the parsing of formulas, their display, and graphical user
  interface management etc. The traditional model of GUI for a big
  family of these tools (implemented in the emacs editor as
  \textsf{ProofGeneral}) is a text editor with three panes: one
  displaying the actual script, the second the state of the current
  proof, and the last the possible messages to the user. One compiles
  a part of the script by moving a cursor forward and backward.
\end{description}

\newpage

\section{Automation in proof assistants}
\label{sec:auto}

The kernel of interactive theorem provers based on type theory, namely
the logic they implement and their typing algorithm, is now a
well-understood part of such systems. But the very fine granularity of
the proofs one can write directly for the kernel to check makes this
act almost impossible for any relatively involved proofs. This led to
the direction of improving the capacity of these systems to receive
partial proofs, possibly with omitted parts or information, and infer
them back fully, to the point where the kernel can check them. Many
different mechanisms fulfill this task on different levels:
\emph{coercions} permit to assimilate notions corresponding to
different underlying objects, \emph{implicit arguments} allow to omit
redundant arguments of functions and theorems, the recent \emph{type
  classes} are an incarnation of the ad-hoc polymorphism present in
some programming languages. To a certain extent, most of the tactics
of \textsf{Matita} or \textsf{Coq} belong to this category as well:
their application triggers the construction of complex proof objects.

In this section, we will focus on one of the most involved of such
mechanism, commonly called \emph{automation} tactics or mechanism. As
opposed to the others, the goal of these is to provide a high-level
method for (more-or-less blindly) searching a proof of a given
statement. The utility of these helpers are needless to be proved
anymore, but we want to advocate here a particular use of them. Two
problems arise from the current, low-level view on formalization.
First, devising proofs directly in type theory or with the help of
low-level tactics is a tedious exercise and reading an already proof
is often difficult. Secondly, the different variations and
peculiarities of each proof assistants make their proofs not
interoperable, resulting in isolated, balkanized formalizations and
duplicated efforts. The development of the automation techniques of
all kind are a direction towards the compensation of both these
problems: they allow to alleviate the proof burden and are a required
step towards the design of very light proof languages that can be
embedded in different foundational dialects and proof assistants, and
is therefore necessary for the design of large, formalized mathematics
repositories.

\subsection{State of the art}

We begin by reviewing some of the common techniques of automated
theorem proving, and discuss their adaptation and limitations in the
setting of type theory, along with implemented generic methods of
proof search in proof assistants.

\subsubsection{First-order theorem proving}

\paragraph{Automated theorem proving}

Automated theorem proving, or automated deduction, is the field
dedicated to the search of proofs of mathematical theorems by computer
software. Theorem provers are tools taking a mathematical statement as
input, and outputting a proof of it (or just succeed) or fail or loop
forever if the statement isn't provable. Most general purpose theorem
provers rely on classical first-order logic (FOL).  These include for
example \textsf{Vampire}, \textsf{Ergo}, \textsf{Waldmeister},
\textsf{Mace4/Prover9}, \textsf{SPASS} or more recently
\textsf{Imogen}. In this part we will review some of the popular
techniques and framework for classical and intuitionistic first-order
theorem proving and discuss their potential use in type theory.

\emph{Resolution} of \cite{robinson1965machine} is the one, if not the
most popular complete method for semi-deciding the refutation proof
problem: given a statement, we can apply this method; if the statement
is unsatisfiable in FOL, the method will eventually terminate with a
refutation of it. The design and implementation of such a method is
discussed at length in \cite{riazanov2002design}.  Many recent provers
rely on resolution, which has proved empirically to be the most
efficient. Since Robinson's discovery, many complete refinements have
been proposed since its introduction: ordered resolution,
selection etc. (see \cite{bachmair2001resolution} for an overview).
It consists of two phases: the transformation of the problem into
clausal form, and the resolution itself, which is the iterated
application of a (single, complete) rule mimicking the \emph{cut} rule
of natural deduction.

Equality is an ubiquitous notion in mathematics. Yet, FOL doesn't
include any special treatment for this symbol, which makes it very
inefficient to use in practice. A better way to treat the equality is
by means of rewriting. It is the goal of \emph{paramodulation}
(\cite{wos1968paramodulation}), \emph{superposition} and their
derivatives (see \cite{nieuwenhuis2001paramodulation} for a complete
survey).  They all are dedicated inference rules for treating equality
externally to the logic, as a mean to replace equals by equals in
formulas. For this purpose, \emph{completion}
(\cite{bachmair2001resolution})is a useful tool to orient equations
and avoid uncontrolled rewriting as much as possible.

Besides these popular techniques, the \emph{tableaux} method and their
variants (see \cite{hahnle2001tableaux}) is a proof search method for
first-order sequent calculus. It has been implemented in various
theorem provers, e.g. 3TAP (\cite{beckert1996tableau}) and the one
from \cite{paulson99}. FOL extended with inductive definitions has
been also the object of studies in ATP systems, resulting in methods
like \emph{rippling} (\cite{bundy1993rippling},
\cite{bundy2001automation}), or \emph{inductionless induction}
(\cite{comon2001inductionless}). The first is a method to cope with
the proofs of inductive cases by syntactic methods, the second is an
efficient embedding of induction into a extension of FOL.
Intuitionistic theorem proving is still an active field of research, a
recent success being \textsf{Imogen} from
\cite{mclaughlin2009efficient}, relying on the focusing discipline
(\cite{andreoli1992logic}).

\subsubsection{Proof search in ITP systems}

In order to give more inference power to existing proof assistants,
recent efforts have been put on adapting the efficient proof-search
methods designed by the automated deduction community to proof
assistants, by adapting them in order to generate proofs and translate
them to the tool's format, for it to recheck them afterward. We will
first do a quick review of the efforts in this domain.

\paragraph{Embedding first-order calculi}

First-order theorem automated theorem provers and the paradigms they
rely on are rich and very efficient. It would seem quite natural to
make use of these advanced tools for interactive theorem proving. One
can achieve this in two ways, either by interfacing an existing tool
through e.g. a dedicated tactic, or by designing a purpose-built
theorem prover that fits the particular needs of the interactive
system. \textsf{PVS} has a rich collection of decision procedures and
generic proof search tools. \textsf{HOL} was interfaced with various
first-order theorem provers such as resolution-based \textsf{Gandalf}
(\cite{hurd1999integrating}) and \textsf{Metis}
(\cite{hurd2003first}). \textsf{Isabelle} was first interfaced with
its dedicated prover based on the tableaux method by \cite{paulson99}
and more recently with \textsf{Vampire} (\cite{meng2006automation}). A
generic communication protocol between \textsf{Coq} and external
provers is being developed (the tactic \textsf{extern}). Since
recently, \textsf{Matita} uses its own paramodulation-based tool
(\cite{asperti2010paramodulation}).

\paragraph{Generic proof search in type theory}

Besides the embedding of first-order automation techniques, proof
assistants generally come with some generic proof search facilities,
devised this time entirely in the type theory. These tactics, known as
\textsf{auto} in \textsf{Coq} and \textsf{//} in \textsf{Matita}, have
at their core a mechanism analogous to the resolution process
introduced earlier. There are important differences with the
integration of external tools as introduced above: their integration
directly into the type theory allows them to exploit its full
higher-order nature, along with their implementation details like
definitions unfolding or inference mechanisms.

\subsection{Issues in ITP automation}

\subsubsection{Embedding first-order calculi}

The deep conceptual differences between both approaches (interactive
and automated theorem proving, higher-order vs. first-order) raises
questions that still need generic answers. We plan on developing the
line of work initiated by \cite{asperti2010paramodulation} in
\textsf{Matita}.

\paragraph{Proof translation}

Some of the interfaces from interactive systems to automatic ones
share the same architecture, at least in systems based on type theory.
We will expose here the approach taken in
\cite{asperti2010paramodulation} With the exception of \textsf{PVS}
which doesn't require a proof and trusts its decision procedures, they
all need to get back a \emph{trace} from the external tool to
reconstruct a proof in their own language. The architecture is then
composed of:
\begin{enumerate}
\item A forgetful translation from the type theory to the
  (first-order) language of the tool. It takes the current goal along
  with a set of hypothesis to a (axioms, problem) pair. This pair is
  run by the external tool and possibly returns a trace of the proof;
\item An interpretation function, which interprets the trace into a
  (possibly incomplete) proof object in the type theory;
\item A retyping mechanism, which infers the missing part of the
  returned proof object and checks it in the type theory.
\end{enumerate}

From this methodology, we see that the integration of external tools
relying on weaker or at least different logics is necessarily an
incomplete method based on heuristics, and it is often difficult to
predict their behavior.

The forgetful translation is not only the operation of removing all
type information from a statement to get a first-order term: type
theory possibly performs computations in statements, has defined
objects that can be unfolded. For example, how do we translate the
higher-order goal $\forall n\ m\ p, (n+1) * m \leq (n+1) * p \to m <
p$? Possibilities are numerous: $S\ n * m \leq S\ n * p \vdash S\
m\leq p$ or $m + n * m \leq p + n * p \to m+1 \leq p$ or any
intermediate steps of unfolding the constants $+$, $*$ or $<$ and
reducing their content. Depending on the set of axioms provided to the
external tool, some will be provable, some not.

The interpretation of trace object is also a delicate operation,
heavily depending on the calculus used by the external tool. Traces
are representations of the success of the proof process: they could be
the list of rewriting steps from the axioms to the problem statement
for paramodulation, or the list of clauses and literals used by the
resolution. The interpretation of a paramodulation step is pretty easy
since its logical meaning (rewriting) is admissible in type theory (by
means of the induction scheme of equality). It becomes quite involved
when considering a \emph{blind} method like resolution: the initial
preprocessing transforming the problem into a set of clause, which is
not admissible in intuitionistic logic, has to be traced carefully and
checked for validity. The same goes for another blind method, SMT
solving.

Finally the last part of the process -- retyping the result of the
interpretation -- is not guaranteed to be feasible since the type
information was lost in the first place. Sometimes the external prover
would success and return a valid trace of the first-order translated
goal, which has no meaning in the type theory and couldn't be
retyped. The inference of the missing type information can also be
quite involved and may itself rely on proof search (see
\cite{asperti2010paramodulation} for details).

\paragraph{Automated vs. interactive paradigms}

But the conceptual differences between automated and interactive
theorem proving go beyond the only untyped vs. typed or first-order
vs. higher-order problems.

One lies in the very purpose of automation in these two paradigms: ATP
systems are usually provided with a small set of axioms for a given
theory, often minimal, and deduce new facts from these onwards, until
the goal is reached. Thus, they generate \emph{cut-free} proofs, in
the sense that they don't infer useful intermediate lemmas that would
need to be proved afterward. On the other hand, automation in ITP
system usually relies on a huge set of objects, both axioms defined by
the user, logical rules in the form of constructors and destructors
for logical connectives, and intermediate proved lemmas, and this set
grows each time a new object is introduced.

From this emerges two difficulties of integrating ITP techniques into
ATP. The first is that provers are not necessarily optimized for
treating big sets of initial axioms, nor for dynamically extending
those sets as is required for an efficient integration. The
pre-treatment of generating and translating these sets might be
prohibitive if not correctly designed in the ITP software from the
beginning. The second and most important difficulty is the one of
choosing among those sets the needed objects. Usually, the whole
library of proof assistants contain very redundant objects: proofs
trivially subsuming other proofs (e.g. by computation, or by
superposition of two results), different proofs of the same statement,
different implementations of the same logical connective\ldots\ To be
treated efficiently and avoiding duplicated work, one would have to
choose in the whole set of objects of the library the relevant subset
for the given problem, minimal but with enough information for the
proof to succeed. This issue has not been throughly tackled and would
need a careful study.

\subsubsection{Generic proof search in type theory}
\label{sec:psitp}

The tactics and methods of proof search as implemented in \textsf{Coq}
and \textsf{Matita} can fulfill different purposes and usages, that we
can classify according to two criteria. First, we discriminate on the
\emph{purpose} intended by the user: it might be used to automatically
fill in some of the small ``trivial'' gaps leaved by the user in a
proof (what we call small-scale automation), for example leaving to
the machine the inference of one trivial case among others; it might
also be intended as a help to devise the proof when the user totally
lacks intuition of it (large-scale automation). This second purpose is
obviously more difficult to achieve, and it is probably unrealistic
considering the results obtained in the dedicated field of automated
theorem proving. Secondly, we separate two different \emph{modus
  operandi} of the automation mechanism: some of them allow to
automatically finish a proof and fill in the proof tree until its
leaves, may it be a trivial step of reasoning (small-scale) or a more
involved one requiring to explore deeply the search space
(large-scale). These take the form of classic tactics like
\textsf{auto} or \textsf{//}. Some of them on the other hand are
\emph{interleaved} in the proof process and allow to incarnate a
certain \emph{quotient} on applicable proof steps or statements. This
is the case of \textsf{Coq}'s type classes, or the recent \emph{smart
  apply} tactic of \textsf{Matita}.

\paragraph{Intuitionistic resolution}

As explained earlier, the resolution mechanism as usually presented is
classical by nature. However, it is well known that this appears as
nothing more than a notational facility, and its core reasoning scheme
is in fact intuitionistic. It proceeds simply by maintaining a list of
hypothesis and lemmas along with the goal; at each step, it applies in
turn all hypothesis and lemmas whose head unifies with the current
goal:
\begin{mathpar}
  \infer[Intro]{\Gamma,x:A\vdash B}{\Gamma\vdash \Pi x:A. B} \and
  \infer[Apply]{\Gamma\vdash A_1\ \ldots\ \Gamma\vdash A_n \and
    \Gamma\vdash B\equiv C \and f:\Pi (x_1:A_1)\ldots(x_n:A_n). B \in
    \Gamma}{\Gamma\vdash C}
\end{mathpar}

This \emph{backward-chaining} method (from the goal to the
hypothesis), although not complete, helps the completion of easy
goals. Because of the dependent nature of the statements in the CIC
however, this process is made more complex by the fact that the
application of a \textsc{Apply} rule can introduce
\emph{metavariables} in the goal, that is holes in a term of a known
type. For example, the application of a transitivity lemma on the
relation $R$ to a goal of the form $R(a,b)$ produces two goals sharing
a metavariable: $R(a,\mathsf{X})$ and $R(\mathsf{X},b)$. This
complicates greatly the control of the proof-search process: one could
na\"ively think of it as a backtracking process on each \textsc{Apply}
rules in turn, but this is erroneous: a goal, even closed by the
application of a 0-ary application (an instance of the \textsc{Init}
rule of natural deduction), could result in a backtracking if a
metavariable was instantiated. Imagine in our example that the goal
$R(a,\mathsf{X})$ is closed by a proof of $R(a,c)$, but that the
second instantiated branch $R(c,b)$ fails to be closed. One has then
to backtrack on the first goal rather than on the previous one, to be
able to find a second, successful instantiation e.g. $R(a,d)$ and
$R(d,b)$.

Due to some limitation of its implemented tactic language, this
subtlety is not taken into account in \textsf{Coq}; whereas it is
solved correctly by \textsf{Matita}. Both implement a depth-first
algorithm of the search tree.

\paragraph{Choice of hypothesis}

Contexts of proof assistants are not simple objects as are their
theoretical counterparts above: they are composed of many segments and
many kind of different objects, like definitions, inductive objects,
constructors, axioms. One question relative to these sort of
automation is: should we consider all of them as equally probable
candidates? Both proof assistants solve the question quite
differently. They both implement a mechanism of priority among these
objects, taking some heuristics to determine them like the number of
goals opened by the application or the number of metavariables
introduced. Besides that, \textsf{Matita} takes however the approach
of considering all objects as candidates except the most general ones
(induction principles for example that can be applied to any goal),
whereas \textsf{Coq} chose the approach of user-defined \emph{hints}
databases, associating to each lemma a priority and a strategy such as
\textsf{Resolve} (general strategy) or \textsf{Immediate} (the goal
has to be closed in one step after its application). Possible
directions like the refinement of these heuristics as well as the fine
control of the strategies are discussed in \ref{sec:strat}.

\paragraph{Definitions and computations}

Due to its higher-order nature and its definitional mechanism, the
logics underlying these proof assistants are subject to some notions
of reduction. Actually, the conversion rule of the CIC ensures that
the statements of the logic are all quotiented w.r.t computation.
These computations can be split into two different forms, that serve
different uses: first, some real computation can occur, as in $P(2+2)
= P(4)$ or $Q(S\ n * m) = Q(n * m + n)$, and a lemma applicable to one
of its part should be equally applicable to the other. The second is
the use of the \emph{definitional mechanism} to abbreviate some
often-used statement forms: for example, one might want to define
$\mathrm{transitive}(R)$ to expand to $\forall xyz, R(x,y) \to
R(y,z)\to R(x,z)$.

This is the responsibility of the \emph{unification} algorithm to
generally determine when these situation happen, when applying a lemma
for example. It is a complex part of a proof system, and often
involves expansive computations. Thus, iterating this process again
and again when trying to \textsc{Apply} all possible lemmas during
proof search can lead to substantial delays. To achieve good
performance, indexing techniques are generally used to rule out as
fast as possible lemmas that don't need to be treated for a given goal
because they won't ever unify. However, these data structures are
purely first-order, and it is non-trivial to use them enough so as to
improve the efficiency but not to restrict too much the candidates and
miss some possibly successful ones. These techniques have been
experimented, and are discussed in more details in
\ref{sec:implem}. In particular, the adaptation of these
data structures to take abbreviation-unfolding is a non-trivial, open
problem.

\subsection{Further directions}

Besides the integration of the first-order prover in \textsf{Matita},
We would like to improve the existing general-purpose proof search
techniques available in these proof assistants. The general
methodology adopted is to design a procedure directly for the Calculus
of Inductive Constructions, by focusing as much as we can on the
user's need. A preliminary analysis has been done by identifying clear
examples of situations where automation is wanted but that the tools
available (in \textsf{Coq} for the moment) are unable to cope with.

\subsubsection{Sequent calculus proof search}

The calculus of constructions is usually implemented in a natural
deduction fashion, at least in \textsf{Coq} and \textsf{Matita}: the
context is used only to retrieve previously defined object, and there
is no way to modify it after its introduction. This proof style is
supported by the \textsf{apply} family of tactics that applies a lemma
or an hypothesis to the goal, progressing in a
\emph{backward-reasoning} fashion: one progresses from the goal to the
assumptions. This approach is however not always intuitive, and a
second kind of reasoning is supported in both tools. The
\textsf{lapply} family of tactics (or \textsf{apply with}) apply a
lemma or an hypothesis to another hypothesis in the dual
\emph{forward-reasoning} fashion. In this case, one progresses by
transforming the assumptions until they eventually match the goal.
These two complementary approaches correspond to two presentation of
the theory of types, respectively the natural deduction and sequent
calculus style. Pure Type Sequent Calculus (PTSC) was introduced in
\cite{lengrand2006normalisation}.

\paragraph{Forward-reasoning}

The automation methods described in \ref{sec:psitp} are, to the best
of our knowledge, relying on a \emph{backward-reasoning} mechanism:
the system applies iteratively all possible applicable lemma and
hypothesis. This method has severe disadvantages: consider the sequent
$A\wedge B\vdash A$. The only way of proving it automatically is to
introduce a metavariable and apply the projection $proj2 : \forall X
Y, X\wedge Y \to X$ to produce the sequent $A\wedge B\vdash A\wedge
\mathsf{Y}$ and to conclude by instantiating \textsf{Y} to
$B$. However, allowing the application of these projections, and most
lemmas that produce metavariables leads to an enormous search space
since they can be applied to any goals. They should be avoided as much
as possible, preferring simpler proofs: here an obvious solution is to
apply the rule of sequent calculus:
$$ \infer{\Gamma, A, B\vdash C}{\Gamma, A\wedge B\vdash C} $$
to obtain the sequent $A,B\vdash A$ and conclude immediately.

In the CIC, the conjunction is a defined object. In fact, $\wedge$ is
an inductively defined non-recursive predicate with one constructor
and two independent hypothesis: $A\wedge B := conj : A \to B \to
A\wedge B$. By $\alpha$-conversion, the above rule of the sequent
calculus is valid if we replace $\wedge$ by any connective of the same
shape. It is also valid for connectives of the same ``kind'': if we call
conjunctions all inductive definitions having an unique constructor,
we can devise rules of forward-reasoning for all of them.
\begin{example} All of $\exists$, $\wedge$ and record types are
  conjunctions. Therefore we can devise the rules:
  \begin{mathpar}
    \infer{\Gamma, A, B\vdash C}{\Gamma, A\wedge B\vdash C} \and
    \infer{\Gamma, x:A, B(x)\vdash C}{\Gamma, \exists x:A. B\vdash C}
    \and \infer{\Gamma, x_1:A_1, x_n:A_n\vdash C}{\Gamma,
      \{x_1:A_1;\dots;x_n:A_n\}\vdash C}
  \end{mathpar}
\end{example}

\paragraph{The case of disjunction}

Can we treat inductive definitions with several constructors the same
way? The na\"ive application of the previous idea leads to a very
redundant proof search. Consider the goal $A\vee B\vdash C$. If we
apply the rule of left disjunction to it, we are left with two goals
$A\vdash C$ and $B\vdash C$. If we can prove the sequent $\vdash \pi :
C$ without using either $A$ or $B$, then the previous application
doubles the work: we have to copy the proof $\pi$ in both
sequent. Therefore, the left rules are to be applied more carefully
for disjunctions.

A second (but wrong) idea is to apply them lazily: we keep them
intact, until the moment where we meet one of their components in the
goal. Then we can destruct them and continue with the second branch:
$$ \infer*{\infer*{\infer*{\infer*{ }{\Gamma, A\vdash A} \and
      \infer*{\vdots}{\Gamma,B\vdash A}}{\Gamma, A\vee B\vdash
      A}}{\vdots}}{\Gamma, A\vee B\vdash C}
$$ Unfortunately this method is not complete, as it may be possible to
prove $\Gamma, B\vdash C$ but not the degraded judgement $\Gamma,
B\vdash A$.

We propose the following improvement: each disjunction sees its
components recorded, and proof search continues leaving the
disjunction untouched. If an instance of one of the component is met
later once a proof $\pi$ has been generated, we backtrack to the point
where the disjunction was introduced, break the disjunction, apply
$\pi$ on its corresponding component, and we are left with the second
branch of the disjunction in the hypothesis.
\begin{mathpar}
  \infer*{\infer*{\infer*{\vdots}{\Gamma,A\vee B\vdash A}}{\pi\left\{\vdots\right.}}{\Gamma,A\vee B\vdash C}
  \and \Longrightarrow \and
  \infer*{\infer*{\infer*{\infer*{ }{\Gamma,A\vdash
          A}}{\left.\vdots\right\}\pi}}{\Gamma,A\vdash C} \and
    \infer*{\vdots}{\Gamma,B\vdash C}}{\Gamma,A\vee B\vdash C}
\end{mathpar}

To sum up, this technique allows to rewrite part of a proof
\emph{during} proof search to optimize its number of inferences. This
way, the depth of a proof can be reduced and since most automated
proof search have their depth bound, we increase the inference power
of these techniques.

\newcommand{\nat}{\ensuremath{\mathbb N}}

Our goal is to generalize these techniques to all ``kinds'' of
inductive. The strategy described above could be generalized to
recursive inductive by taking into account the induction hypothesis,
then performing induction by itself:
\begin{mathpar}
  \infer*{\infer*{\infer*{\vdots}{\Gamma,n:\nat\vdash
        C(0)}}{\pi\left\{\vdots\right.}}{\Gamma,n:\nat\vdash C(n)}
  \and \Longrightarrow \and \infer*{\infer*{\infer*{\infer*{
        }{\Gamma,C(0)\vdash
          C(0)}}{\left.\vdots\right\}\pi}}{\Gamma\vdash C(0)} \and
    \infer*{\vdots}{\Gamma,m:\nat,C(m)\vdash
      C(\textsf{S}(m))}}{\Gamma,n:\nat\vdash C}
\end{mathpar}

By combining forward- and backward-reasoning step in an efficient way
and respecting the underlying logic, we believe that we can build an
automatic proof search method, not only more efficient than the
currently implemented methods, but also more expressive and able to
cope with the problem of deferring ``trivial'' proof steps to the
theorem prover, leaving to the user the load of only sketching the
proof.

\subsubsection{Strategic proof search}
\label{sec:strat}

Proof search however is not only about data structures and inference
rules. It is a dynamic process involving complex backtracking among
goals, and even more complex in the presence of metavariables. The
goal of fast automation requires a fine control on backtracking on the
search tree, as was already hinted in the previous section: attaining
a given goal made us backtrack to a special position and remember some
pieces of information (the proof $\pi$) to replay it without possible
backtracking on the next goal.

\paragraph{Focusing}

We can go further in this idea by putting into practice ideas that
emerged in the \emph{focusing} discipline of \cite{andreoli1992logic}.
Focusing was introduced as a winning strategy for proof search in the
Linear Logic of \cite{girard1995linear}, exploiting the redundancies
present in na\"ive proof search. These redundancies are actually not
on space, but on time, i.e. it is not a problem of data structure for
the proofs (as was solved by \emph{proof nets}), but a problem of
backtracking in the proof search. Indeed, it was identified that some
rules of Linear Logic are \emph{invertible}, that is, they can be read
in both directions, from top to bottom or from bottom to top. Thus the
application of such a rule does not change the provability of a
sequent and can be applied \emph{eagerly}. The eager application of a
rule is its application without the need to backtrack before them:
these rules just simplify the goal without modifying their
provability. Examples of invertible rules are the rules on conjunction
and disjunction, both left and right. This remark led to the
specification of \emph{polarities} defined on sub-formulas: the
\emph{negative} parts of a formula comprises all chain of invertible
connectives, whereas the \emph{positive} parts make up for the
non-invertible one. A \emph{phase change} occurs during proof search
when a connective is met that is not of the same polarity as the
previous one. Focusing is the strategy defined by:
\begin{itemize}
\item Applying eagerly all the connectives during the negative phase,
\item Focusing on an arbitrary hypothesis during the positive phase
  and apply eagerly all possible rules,
\item Backtracking occurs only at phase change.
\end{itemize}

These ideas, although having been defined for linear logic, can apply
to intuitionistic logic as well, by embedding one logic into another
by a (non-unique) polarized translation. The automated theorem prover
\textsf{Imogen} by \cite{mclaughlin2009efficient} is based on these
ideas, and proved to be a success.

A focusing approach applied to dependent type theory is still to
devise though, and would constitute a theoretical advance with clear
applications to automated proof search in type theory.

\paragraph{Backtracking strategies}

More generally, an efficient proof search requires a control of
backtracking that goes beyond what is currently implemented in
\textsf{Coq} and \textsf{Matita}: both proof assistants consider the
all leaf of the search tree as potential backtrack nodes, and don't
provide a way to specify strategies in a general way. Furthermore, no
distinction is made in the search between logical objects like the
defined connectives or the equality, and full lemmas with non-trivial
proofs. \textsf{Coq}'s $\mathcal L_{tac}$
(\cite{delahaye2001conception}) is an attempt to provide a
full-featured programming language for automatic tactic application.
It provides a recursion operator as well as a matching construct on
the current goal. However, it fails to define general and extensible
way of defining general-purpose automation tactics, specifically to
deal with non-determinism, and is confined to domain-specific
application, like the propositional fragment of logic
(\textsf{tauto}), or the decision of some order-related statements
(\textsf{order}).

We propose, as a first step towards the design of strategies, to have
a clear developer-side language of specification of these strategies,
expressive enough to cope with previous ideas: backtracking by keeping
a (partial) proof, focusing on a particular formula, eagerly advance
on the application of a particular rule. Also, we would like to
abstract from the \emph{names} of the inductive objects to select them
according to their \emph{kind}: (dependent, indexed, recursive)
disjunctions, conjunctions\ldots\ The design of such a general method
is of course interleaved with, and will be nurtured by the discovery
of particular wanted strategies. A good starting point could be the
backtracking monad of \cite{kiselyov2005backtracking}.

\subsubsection{Implementation issues}
\label{sec:implem}

When it comes to implementation, all the techniques mentioned above
require the manipulation of large sets of terms, and a recurring
problem is to retrieve terms satisfying some structural conditions,
like unification with a given pattern. One-to-one unification has a
well-known, simple and efficient algorithm, but it becomes
under-optimal when applied iteratively on large sets of terms, because
a lot of sharing is lost. Term indexes are data structure used to
store these sets, making the retrieval of unifiers to a given pattern
more efficient.

One data structure used for this purpose has been \emph{discrimination
  trees} or \emph{nets} (\cite{mccune1992experiments}). It operates on
the same principle as \emph{tries}, viewing the terms as flattened
strings and retrieving following a given prefix (the pattern), like a
dictionary. A critical inefficiency of this structure is due to the
loss of information resulting from the flattening of
terms. \emph{Substitution trees} (\cite{graf1995substitution}) are a
popular alternative. Their main idea is to index the terms according
to their \emph{mgu}'s, and form a tree of substitutions. All these
techniques are devised for first-order term algebras. With the
emergence of automated provers for higher-order logics like
\textsf{Twelf}, some attention has been drawn on the problem of
\emph{higher order indexing}, i.e. indexing of $\lambda$-terms modulo
$\beta$-reduction (\cite{pientka2003higher}).

This problem is in general undecidable, but such a behavior is
definitely a wanted feature for our use. We will focus at first on a
subproblem, already interesting and not yet tackled: the goal of
adapting these structure to be used modulo definitional equality to
abstract from certain abbreviations used commonly.

Moreover, if we decide to tackle the problem of the implementation of
sequent calculus proof search in a proof assistant, where the search
is not only goal-directed anymore but to some extend also directed by
the hypotheses, we then want to index hypothesis and be able to
retrieve efficiently a term having e.g. a given form of hypothesis.
However, due to the set nature of the context, the pair (hypothesis,
goal) cannot be presented as a term since it is to a certain extend
possible to permute all hypothesis. We then need to devise an
efficient data structure of term indexing taking into account to these
permutations. It seems not trivial to adapt existing term indexing
methods for this purpose. Moreover, hypothesis in a context are in
type theory possibly dependent: the instantiation of one hypothesis
triggers the modification of all subsequent dependent hypothesis. Our
data structure should be able to instantiate the variables
accordingly. This is however a known problem, and most of the time a
simple solution exists: the carry of a closure substitution during the
instantiation.

\subsubsection{Isomorphisms in the CIC}

The calculus of inductive constructions has the nice property of being
a very minimal foundational language, in which even very low-level
mathematical concepts can be defined. For example, the usual logical
connectors $\vee, \exists, \wedge\ldots$, even the equality are not
elementary objects but defined ones, thanks to the dependent and
inductive types features. However, this freedom implies that a given
usual \emph{concept} can have a lot of different
\emph{implementations}. Some of them will be convenient or efficient
for a given task, other for other tasks.

A simple example is the definition of the logical equivalence in
\textsf{Coq}'s standard library. It is historically defined as
$A\Leftrightarrow B := A\to B \land B\to A$, but a recent idea was to
change it to an inductive definition $A\Leftrightarrow B :=
\mathsf{build\_iff} : (A\to B)\to(B\to A)\to A\Leftrightarrow
B$. Although both definitions are strictly equivalent and represent
the same ``usual'' notion, such a change implied to rewrite almost all
proofs relying on the first definition and was then aborted.

Another less-trivial example is the representation of (natural)
numbers. They accept a wealth of different structural definitions ---
Peano, binary, machine, axiomatic integers\ldots --- each supporting
its own reasoning style, having different efficiency properties
etc. but, in the end, representing the same usual notion. When one
wants to begin a new development using natural numbers, she has to
choose from the beginning her implementation and stick to it
throughout the development. A change of representation would require
to rewrite a large part of the development and is therefore not
affordable; moreover, accessing the results, or using properties of
another representation is simply forbidden.

These facts suggest that a certain notion of \emph{isomorphism}, or
simply of \emph{morphisms} between the \emph{concepts} to be
formalized and their (possibly numerous) \emph{implementation} would
be a valuable study for both the modularity of the proofs (more
results to share among developments) and the abstraction of the proof
language, which would rely not on the actual implementation details of
structures but on higher-level concepts. The notion of type
isomorphisms has been already well-studied in the setting of
functional programming languages like ML, notably by
\cite{di1995isomorphisms} and proposals of integration in type theory
by \cite{barthe2001type} have been made.

There seem to be many different approaches to this problem, some
already considered, some not. A first approach is the definition of
axiomatic \emph{interfaces} for data structures or concepts, by means
of records or module system in the fashion of modular programming,
like it is done in the formalizations of algebra undertaken in the
CIC. A natural number is a ``\emph{complete}'' set of its
``\emph{atomic}'' properties (terms to be defined): successor lemmas,
recursion\ldots\ Individual representations of natural numbers are
then implementations of the interface. But what are these properties,
and how can one be sure that they are ``minimal'' or ``complete''
wrt. the common concepts, or appropriate for her particular needs?
Another pragmatic approach is to ``elect'' a canonical representation,
and to have translation functions/morphisms from each of the other
representations to it. This way, we lose their interesting structural
properties (e.g. efficiency), but we inherit from all properties
proved on them. A more \emph{bottom-up} but automatizable proposition
is to work directly at the syntactic level and assimilate isomorphic
terms in the calculus. Contrary to the other suggestions, this is a
meta-investigation of the isomorphisms in the logical framework that
could lead to the (partial) automation of reasoning modulo some
implementational details (like e.g. the \emph{iff} example
above). Inductive types seem to provide a rich framework for studying
these isomorphisms, since their purpose is precisely to provide
one-to-one correspondence between a ``concept'' (the inductive type)
and its possible alternative ``implementations'' (the
constructors). For example, we could assimilate function types which
are equal modulo permutation of arguments, or inductive types sharing
the same structure, types defined on an inductive type and the
corresponding substituted inductive type\ldots\ Building up on this
principle, we can hope to capture a large class of equivalences. A
third proposition is to rely on existing automation techniques to deal
with isomorphisms. Most instances of these isomorphisms can be
expressed as equalities, for example curryfying propositions: $\forall
A B C, A\to(B\to C) = (A\wedge B)\to C$. By providing a way to apply
logical rules up-to these equalities, we can somehow abstract from a
part of the logical specificities. This is the approach taken in the
new \texttt{sapply} facility of \textsf{Matita} described in
\cite{asperti2010paramodulation}.

The applications of such a study are numerous: abstract proof and
logical language mapped ``by construction'' to the foundational
language, modularity of the developments, increased automation\ldots

% \subsubsection{Proof mining}

% Over the years, large developments of formally verified proofs have
% been made in various fields of science, from mathematics to hardware
% verification, with the help of various interactive proof
% assistants. These large proofs along with many others are stored in
% repositories and can be rechecked at will. Yet, the information
% provided by these developments remains largely unexploited. Most proof
% assistants provide a very minimal set of interaction with previous
% developments: loading, reusing a result, searching by statement, and
% usually some statement-based automation like described above. In these
% interactions, the proofs themselves are totally passive, kept only to
% testify of the validity of the reasoning. Unlike their equivalents in
% programming languages---libraries---proof assistants repositories do
% not only consist of names of objects (functions, classes etc.)  but
% they also form a very structured and valuable source of information,
% both on the proof process and on the use one makes of the logical
% framework.

% By analyzing these large databases of already human-formalized proofs
% with the help of \emph{empirical methods} (statistics, data-mining,
% information theory, machine-learning), one could extract some useful
% information. It is expected to be of interest for a number of
% tasks. By analyzing them ``by hand'', some common reasoning schemes
% could be isolated and serve as a basis for improving or designing
% high-level proof languages and procedures, validating their design
% \emph{by the usage}.  It could also, as a side-effect, be treated
% automatically or semi-automatically, to provide a valuable help to the
% user in a given situation about what has been previously done in
% similar situations -- taking the form of suggestions or hints, or even
% to fully automatize some easy parts of proofs in a machine-learning
% fashion. The analysis of proof patterns could help also to
% automatically factorize large proofs, to infer useful intermediate
% lemmas or to compress the representation of a proof. This methodology
% has already been advocated and experimented in \cite{duncan2004udm}
% with the help of standard data-mining concepts, and \cite{urban08} in
% a machine-learning fashion, both leading to some interesting
% results. Their approaches however, concentrates on some features of
% the proofs. For example, they analyze the uses of lemmas and axioms in
% a given proof, preferring in a future situation a more productive
% lemma over one that hasn't been used at all. Proofs are complex
% objects, which structure itself is much richer than these features,
% and it might be even more productive to reflect the structure itself
% in the analysis.

% Among all information contained in the databases of formal proofs
% consisting proof assistants' standard libraries, the order of chaining
% of logical rules in the proofs, i.e. their very \emph{structure}, is
% already of crucial interest: it could contain enough information to
% both spot previously unknown practices of mathematicians, and even
% help the resolution of analogous cases, as an automation procedure. By
% extracting from these big databases of proof trees the common,
% recurring or generally interesting chaining of rules, that is some
% sub-trees of our trees, we can expect to isolate some information
% about the proof process, even some useful methods for future proofs.

% We propose to data-mine proofs by means of \emph{frequent subtree
%   mining}. Frequent subtree mining is a relatively new research field,
% dedicated to the extraction of recurring patterns inside the structure
% of trees (see an overview in \cite{chi2005frequent}). These methods
% will have to be adapted for our purpose: proof-terms have a certain
% structure that follows from the logical framework they were written
% in, whereas most algorithms in the literature have been devised for
% general trees. The direct reuse of a known algorithm has been proved
% in a preliminary investigation to be unadapted: the large majority of
% results are be non-relevant and the interesting ones drowned in this
% mass. Also, some of the notions used in the algorithms will have to be
% redefined: the notion of subtree differs in presence of (proof-)terms,
% and we will have to refine the definition of frequency to capture what
% are ``interesting'' patterns. Some preliminary results have been
% exposed by the author in \cite{puech2009formal}.

\newpage

\section{Semantic patches}

The second, complementary axis of this proposal is the attempt to
tackle the problem of modularity and evolution in formally checked
mathematical developments.

Even beside the notion of constructivism and the assimilation of
proving and programming in type theory, it is interesting to
investigate the possible relationship between the very
\emph{methodologies} employed in both fields, mathematics and computer
science. And by that we even intend the comparison of the \emph{daily
  workflow} of both scientist: one trying to prove a theorem, the
other constructing a program to fulfill a task. How does one or the
other elaborates his object of study? What kind of \emph{a posteriori}
modification is he prone to doing? What do these modifications imply
on the validity of the whole edifice? How to relate how both
scientists collaborate in a team? How do they rely on existing work to
build up new results?

Recent efforts in the formalization of mathematical results have
naturally led to these questions and many of them remain largely
unanswered, but the tendency seems to be to adapt existing methods
coming from software development, witness for example the recent
introduction of modules or separate compilation in proof assistants
like \Coq or \Matita, the use of dependency management tools
(\textsf{make}) or version control system (\textsf{git}, \textsf{svn})
to build and manage versions of a project. The use of software
engineering techniques for the management of a large formalization
project is even the thesis advocated in the manifesto of the
Mathematical Components team (\cite{gonthiermanifesto}).
\\[-0.5em]

We propose to address a small part of these questions, namely the
enhancement and adaptation of version control systems to the
management of mathematical repositories, by means of what we
preliminarily call a \emph{semantic patch system}\footnote{Joint work
  with \href{http://www.pps.jussieu.fr/~yrg/}{Yann R\'egis-Gianas}}.
It is a generic system allowing one to express \emph{transformations}
on formal developments, and check that some semantic properties are
preserved by the transformations, for example well-typing or
operational equivalence. Eventually, we hope to be capable of
expressing complex transformations on formal proofs, check them
efficiently, and manage distributed repositories of mathematics with
it, guaranteeing \emph{by typing} the stepwise global consistency of
the repository.

In the light of the Curry-Howard isomorphism though, we abstract
totally from whether we are talking about proofs or programs. That is
why, in the following, we start from today's situation, where existing
version control system are used both in software and proof
development, and we will not make any difference between both
use. Depending on his interest, the reader can thus choose his own
side of the isomorphism by translating program into proof, type into
statement\ldots\ or vice versa.

\subsection{Motivations}

The programmer and the mathematician know very well that developing a
new theory or program is not a linear task: one do not open her text
editor and write down her ideas in a unique pass. In the case of
program development, most of the time is actually spent editing
previously written code, correcting bugs or implementing new features
which are interleaved with previous ones. This idea is not present at
all in the text-editor/compiler paradigm, but to reflect it,
\emph{version control systems} usually rely on a textual
differentiation mechanism (\emph{diff}s): the comparison between two
versions of a text file generates a \emph{patch} which indicates which
lines of text are to be added or deleted to transform one into the
other.

Developments, may it be proofs or programs, are usually split into
files, each of these containing a self-contained module of the whole
development, an ``atom'' in some sense. To compile -- or check -- the
whole developments, we use dependency management tools like
\texttt{make}. These tools generate dependencies among the files, and
launch the compilation only on files that have been changed since last
compilation, and their dependencies. This process, known as
\emph{separate compilation} has files as atomic objects, and
dependency generation is performed uniquely on them, ignoring their
internal structure.

As intelligent and widely adopted as these tools have gotten, we
believe that they are not adapted for proof developments. Indeed,
whereas compilation of a program is usually fast enough for the
programmer to rely on the usual interaction loop ((edit; compile)*;
commit)*, the operation of proof checking is usually too expensive
computationally to mimic this workflow. But even besides the time
factor, this ``traditional'' way of formalizing mathematics hinders
the process of mathematical discovery process: once a concept
contained in a file is compiled, it is considered frozen and any
changes to it require the recompilation of all files depending on it;
the linearity of the development also gives no room for alternate,
inequivalent definitions. This fact has nonetheless been shown to be
of crucial to the mathematical discovery process by
\cite{lakatos1964proofs}, and we believe that they should be taken
into account in the realization of mathematical assistants like
\textsf{Coq} or \textsf{Matita}.

Actually, we even dare to state that these tools and the workflow they
suggest were not even adapted for program development in the first
place, but were just a convenient approximation of the user's intent.
Here are some hints to justify this argument: first of all, \emph{any}
change in a source file requires full recompilation of all files
depending on it, whereas a finer management of dependencies is
possible (function-wise for example). The basic operations of a
version control system such as \texttt{svn} or \texttt{git} are often
very sensitive, error-prone and based on heuristics (example, a
\emph{merge}). This is among other due to the fact that the
signal-to-noise ratio of a textual \emph{diff} is often high, because
semantically void changes like altering indentation have the same
value as changing the argument of a function. Finally, as they are not
aware of the meaning of the text they are managing, these tools don't
provide any consistency guarantee on the resulting files (except
textual ones).

All these facts, if they are not a major drawbacks when applied to
weakly structured languages (e.g. assembler, or a \LaTeX\ document),
gain a crucial importance when considering a typed language (a
functional programming language for example) : we would like to make
sure that a patch does not break the well-typing (or any other
property) of a source code.  One more step forward, more
structure-awareness becomes \emph{essential} when they apply to a
proof language, where the structure (the typing) \emph{is} the only
valuable information.

\subsection{Methodology}

We propose here to devise a system for \emph{semantic patches}. It
substitutes the idea of textual transformation by:
\begin{itemize}
\item First preferring an abstract syntax tree representation instead
  of plain text;
\item Secondly embedding semantic data into the transformations, in
  order to be able to reason on them.
\end{itemize}

We want to design a \emph{language of patches}, able to represent
these transformations and their semantic properties, thus capturing
local changes in programs, as well as their global effect on the whole
project. We can see this goal as a refinement of the former idea of
``dependency''.

A semantic patch is a program intended to transform a program written
in an \emph{object language} $\mathcal L$. We focus, on the first
iteration of the project, on the semantic assertion of well-typing:
from the object language's syntax and typing rules $\mathcal L$, we
derive typing rules for the patch languages $\mathcal P(\mathcal L)$,
and we can then type the patch. Our motto is:

\begin{quote} \label{motto}
  A well-typed patch is the guarantee that it transforms a well-typed
  program into a well-typed program.
\end{quote}

In this sense, it is closely related to \emph{incremental typing}: if
a modification has been made on a program, it is expressible by a
patch. To ensure that it will result in a valid program, it suffices
to type the patch, and not the whole resulting program.

Moreover, our approach is meant to be totally formal, so that we can
prove not only validity of our transformations, but also the
completeness of our patch language and other properties. For this, we
place ourselves in a constructive, higher-order metalanguage, namely
the Calculus of Inductive Constructions (CIC), in which we will
formalize in turn:
\begin{enumerate}
\item The object language $\mathcal L$ of our choice
\item Its metatheory
\item The patch language $\mathcal P(\mathcal L)$, parameterized by
  $\mathcal L$.
\item Its metatheory
\end{enumerate}

\subsection{Related work}

Both the study of metatheorical properties as proof transformations
and of change impact in structured documents (e.g. proofs, programs)
is not a new subject.

The \textsf{Twelf} project (\cite{pfenning1999system}) is an
implementation of the Logical Framework (LF,
\cite{harper1993framework}). It is a logic programming language able
to represent logics or programming languages, and embeds an inductive
metatheorem prover able to construct transformations of the
$\Pi\Sigma$ form. It was used in \cite{anderson1993program} to devise
transformations of proofs in order to extract efficient programs.

The problems of managing a formal mathematical library have been dealt
with in various proof assistant and mathematical repositories. The
HELM project (\cite{asperti2000content}, \cite{asperti2006content})
was an attempt to create a large library of mathematics, importing
\textsf{Coq}'s developments into a searchable and browsable database.
Most ideas from this project were imported into the \textsf{Matita}
proof assistants (\cite{asperti2007hop}), especially a mechanism of
\emph{invalidation and regeneration} to ensure the global consistency
of its library w.r.t changes, with granularity the whole definitions
or proofs and their dependencies. The MBase project
(\cite{kohlhase2001mbase}) attempts at creating a web-based,
distributed mathematical knowledge database putting forward the idea
of \emph{development graph} (\cite{hutter2000management},
\cite{autexier2000towards}) to manage changes in the database,
allowing semantic-based retrieval and object-level dependency
management.

This idea, generalized over structured, semi-formal documents gave
birth to \texttt{\it locutor} (\cite{muller2008fine}), a fine-grained
extension of the \texttt{svn} version control system for XML
documents, embedding ontology-driven, user-defined semantic knowledge
which allows to go across the filesystem border. It embeds a
\emph{diff} algorithm, operating on the source text modulo some
equality theory to quotient the syntax. On the same line of work, we
should mention the \emph{Coccinelle} tool
(\cite{padioleau2008documenting}). It is an evolution over textual
patches, specialized on the C language, allowing more flexibility in
the matching process, and was developed to deal with the problem of
\emph{collateral evolutions} in the Linux kernel. It embeds a
declarative language for matching and transforming C source code,
operating on text modulo defined isomorphisms.

Our approach to the ``impact of changes'' problem seems novel on
several aspects: first, it applies uniformly on proofs and programming
languages by virtue of the Curry-Howard isomorphism, and because we
operate at the AST level. Secondly, by taking \emph{types} as a
witnesses for the evolution of a development, we refine the usual,
dependency-based approach for a finer granularity. Thirdly, the
formality of our approach ensures not only a maximal trust in the
system, but also the ability to ``meta-reason'' on the defined
language.
\\

In the following, we will give preliminary hints on the construction
of such a system, and how it could be useful to supersede not only
textual \emph{diff}s and version control systems, but also other
paradigms for modular programming and go towards an eased mathematical
discovery process with the help of proof assistants.

\subsection{Patches as metatheorem composition}

The approach we propose is the following: to transform a program into
another and preserve well-typing, we need to transform accordingly the
typing derivation of the source program into a valid derivation for
the target program. On one hand, these transformations, if performed
by hand, are very error-prone: the theory of the object language may
be tricky and side conditions are frequently omitted in informal
presentations. On the other hand, there is a wealth of formalization
techniques of programming languages in the literature, and numerous
implementations in proof assistants. Theoretical results on the object
language are called \emph{metatheorems}; if these metatheorems are
proved in a constructive formalism, like the one of the proof
assistants \textsf{Coq} or \textsf{Matita}, they constitute algorithms
of program transformation. Then we just have to choose an adequate,
``complete'' set among those. Take the simply typed $\lambda$-calculus
(STLC) for instance. Examples of metatheorems are:
\begin{itemize}
\item \textsf{app} : from two terms $t$ and $u$, we can always form
  the application $tu$. This is a purely syntactical constructor of
  the object language, promoted as a metatheorem;
\item \textsf{typ\_app} : if $\Gamma \vdash t : A \to B$ and $\Gamma
  \vdash u : A$ then $\Gamma\vdash tu : B$. It is part of the typing
  definition of the language, and its proof constitutes a
  transformation building an application node at the root of a
  program;
\item \textsf{inv\_app} : if $\Gamma \vdash tu : B$ then there is a
  type $A$ such that $\Gamma\vdash t:A\to B$ and $\Gamma\vdash
  u:A$. Its proof is a transformation decomposing an application into
  its two subprograms: the function and its argument;
\item \textsf{weaken} : if $\Gamma\vdash t : A$ and $x$ is not free in
  $t$ then $\Gamma, x:B\vdash t:A$. Its proof has no effect on the
  program itself, but enlarges the context in which it is typed.
\item \textsf{strengthen} : if $\Gamma,x:B\vdash t:A$ and $x$ is not
  free in $t$ then $\Gamma\vdash t:A$.  \\\ldots
\end{itemize}
Considering a formalization working directly on the abstract syntax
tree of the programming language (named, possibly implicitly
typed\ldots), our patch language is formed on the composition of a
given, carefully chosen set of metatheorems.

\subsubsection{Definitions}

\paragraph{Transformers}

We consider only a subset of the possible metatheorems, those of the
following form. A \emph{patch variable} is an element $x,y,\ldots$
taken in a countable, infinite set $\mathcal X$. An \emph{atom} is
either a syntactic category of the object language, or a meta property
we want to reason on, i.e. for STLC:
\begin{align*}
  A,B\ ::=\ &\mathrm{var}\ |\ \mathrm{term}\ |\ \mathrm{env}\ |\
  \mathrm{type}\\
  |\ &{}\mathrm{deriv}\ {of}\ (\mathcal X \times \mathcal
  X\times\mathcal X)\\
  |\ &{}\mathrm{variable}\ of\ (\mathcal X\times
  \mathcal X\times\mathcal X) \\
  |\ &{}\ldots
\end{align*}

A \emph{judgement} $j$ can be of two kinds: $x$ is an atom $A$, or $x$
has form $\mathsf{T}(x_1,\ldots,x_n)$, where \textsf{T} is a transformer.
\[ j,k\ ::=\ x:A\ |\ x=\mathsf{T}(x_1,\ldots,x_n)\]

Now a \emph{transformer} $\emph{T}$ corresponds to a metatheorem in
$\Pi\Sigma$ form, with added definitional constraints. The statement
of the metatheorem is ``reified'' into a syntactical \emph{signature}
of the form:
\[ \mathsf{T}\ ::\ j_1, \ldots, j_n \longrightarrow k_1,\ldots, k_m \]

The semantics of these signature is: all tuples on the left of the
arrow are arguments, all tuples on the right are results. Arguments of
kind $x:A$ are interpreted as bare arguments to the transformer,
arguments of kind $x=\mathsf{T}(\vec x)$ as \emph{constraining
  arguments}: the actual argument supplied must be the result of
applying $\mathsf{T}$ to $\vec x$. Dually, results of kind $x:A$ are
bare results of unknown form, whereas results of kind
$x=\mathsf{T}(\vec x)$ are results ensured to be constructed from the
application of $\mathsf{T}$.

\begin{example}
Here are the signatures of some of the metatheorems given above:
\begin{align*}
  \mathsf{app} :: {} & (t\ u : \mathrm{term}) \longrightarrow
  (v : \mathrm{term}) \\
  \mathsf{typ\_app} ::{}& (\Gamma:\mathrm{env}) (A\ B:\mathrm{type})
  (C=\mathsf{arrow}(A,B)) (t\ u:\mathrm{term}) \\
  {}& (d_1 : \mathrm{deriv}(\Gamma,t,C)) (d_2 :
  \mathrm{deriv}(\Gamma,u,A)) \longrightarrow (v=\mathsf{app}(t,u))
  (d:\mathrm{deriv}(\Gamma,v,B)) \\
  \mathsf{inv\_app} ::{}& (\Gamma:\mathrm{env}) (t\ u:\mathrm{term})
  (v=\mathsf{app}(t,u))(d:\mathrm{deriv}(\Gamma,v,B)) \longrightarrow
  \\ {}&
  (A:\mathrm{type})(C=\mathrm{arrow}(A,B))(d_1:\mathrm{deriv}(\Gamma,t,C))
  (d_2:\mathrm{deriv}(\Gamma,u,A))
\end{align*}
but we can also imagine higher-level transformers:
\begin{align*}
  \mathsf{cps} ::{}& (t:\mathrm{term}) \longrightarrow (u:\mathrm{term}) \\
  \mathsf{aa\_trans} ::{}& (A:\mathrm{type}) \longrightarrow
  (B:\mathrm{type}) \\
  \mathsf{typ\_cps} ::{}& (\Gamma:\mathrm{env}) (t:\mathrm{term}) 
  (A:\mathrm{type}) (d:\mathrm{deriv}(\Gamma,t,A)) 
  \longrightarrow \\ {}&
  (t'=\mathsf{cps}(t))
  (A'=\mathsf{aa\_trans}(A)) (d':\mathrm{deriv}(\Gamma,t',A'))
\end{align*}
\end{example}

\paragraph{Patches}

Given a set of transformers $\Sigma$ and their signatures, a
\emph{patch} $\Delta$ is a sequence of variable assignment constructed
from the application of a transformer $\mathsf{T}_i\in\Sigma$ to bound
variables, much like a \textsf{let} construct:
\begin{align*}
  \Delta\ {}::=& \vec x_1 = \mathsf{T_1}(\vec y_1);\ \ldots;
  \\ {}& \vec x_n = \mathsf{T_n}(\vec y_n)
\end{align*}

Here, all $\vec x_i$ are bound in subsequent assignments.  A
\emph{repository} is a patch with no free variable. A patch therefore
represents the evolution of a set of atoms: the free variables
represent what the patch expects as input, all bound variables the
newly constructed objects. A repository represents the history of
modification of a program, starting from ground up. Each variable has
an implicit type, taken into the set of atoms.

We now describe informally a typing discipline for the patches: a
transformer should be applied to exactly its number and type of
arguments, and bind exactly its number of results, which take their
type from the signature. Results of kind $x=\mathsf{T}(\vec y)$ are to
be interpreted as additional assignments specifying the structure of
the variable bound. A transformer with an argument of kind
$x=\mathsf{T}(\vec y)$ should be applied to a variable having exactly
been created by the definiens $\mathsf{T}(\vec y)$: same transformer
name \textsf{T}, same variable names $\vec y$.

As we said earlier, a well-typed repository (in the way we just
hinted) describes well-typed changes of its objects, for examples
programs and derivations. An important property of this typing
discipline is its independence w.r.t the metatheorems associated to
the transformers in $\Sigma$. All the typing process is done
\emph{entirely syntactically}, without having to peak into the
semantics of transformers. This allows the patch language to be only
parametric on $\Sigma$.

\newcommand{\interp}[1]{\llbracket{#1}\rrbracket}

\paragraph{Interpretation}

The same way we reified our metatheorem's types into a syntax (the
transformers) to obtain the patch language, we can do the opposite
transformation, i.e. interpret a patch or a repository $\Delta$ into
an object of the language. This is done the obvious way, by following
the definitions in $\Delta$ and interpreting transformers by their
associated metatheorem:
\begin{align*}
  \interp x_\Delta &= \Delta(x) \\
  \interp{\mathsf{T}(\vec y)} &= T(\interp{\vec y})
\end{align*}

The following result ensures that our motto (\autoref{motto}) is correct:
\begin{theorem}[Soundness]
    If $\Delta$ is well-typed and $d : \mathrm{deriv}(\Gamma,t,A) \in
    \Delta$, then $\interp d_\Delta$ is a derivation of
    $\interp\Gamma_\Delta\vdash\interp t_\Delta : \interp{A}_\Delta $
\end{theorem}

\subsubsection{Examples}

We make a pause here and carry on a simple example of repository
construction, assuming the object language to be the simply typed
$\lambda$-calculus. The transformers involved are the following:

\begin{align*}
  \mathsf{int}::{}&\mathrm{type} \\
  \mathsf{string}::{}&\mathrm{string}\To\mathrm{var}\\
  \mathsf{var}::{}&\mathrm{var} \longrightarrow \mathrm{term} \\
  \mathsf{nil}::{}&\mathrm{env} \\
  \mathsf{cons}::{}&(x:\mathrm{var})(A:\mathrm{type})(\Gamma:\mathrm{env})
  \To
  \mathrm{env} \\
  \mathsf{init}::{}&(\Gamma:\mathrm{env}) (x:\mathrm{var})
  (t=\mathsf{var}(x)) (A:\mathrm{type}) \\ {}&
  (v:\mathrm{variable}(\Gamma,x,A)) \To
  (d:\mathrm{deriv}(\Gamma,t,A))
  \\
  \mathsf{zero}::{}& (\Gamma:\mathrm{env})(x:\mathrm{var})
  (A:\mathrm{type})(\Gamma'=\mathsf{cons}(\Gamma,x,A)) \To \\ {}&
  (v:\mathrm{variable}(\Gamma',x,A)) \\
  \mathsf{succ}::{}& (\Gamma:\mathrm{env}) (x\ y:\mathrm{var}) (A\
  B:\mathrm{type}) (v:\mathrm{variable}(\Gamma,x,A)) \\ {}&
  (\Gamma'=\mathsf{cons}(\Gamma,y,B)) \To
  (v':\mathrm{variable}(\Gamma',x,A)) \\
  \mathsf{typ\_lam}::{}& (\Gamma:\mathrm{env})(x:\mathrm{var})(A\
  B:\mathrm{type}) (t:\mathrm{term}) \\ {}&
  (\Gamma'=\mathsf{cons}(x,A,\Gamma))(d:\mathrm{deriv}(\Gamma',t,B))\To
  \\ {}&
  (t'=\mathsf{lam}(t,x,A))(C=\mathsf{arrow}(A,B))(d':\mathrm{deriv}
  (\Gamma,t',C)) \\
  \mathsf{typ\_app}::{}&(\Gamma:\mathrm{env})(t\ u:\mathrm{term})(A\
  B:\mathrm{type}) (C=\mathsf{arrow}(A,B)) \\ {}&
  (d_1:\mathrm{deriv}(\Gamma,t,C))(d_2:\mathrm{deriv}(\Gamma,u,A))
  \To (d_3:\mathrm{deriv}(\Gamma,v,B)) \\
  \mathsf{weak}::{}&(\Gamma:\mathrm{env})(x:\mathrm{var})(A\
  B:\mathrm{type})(t:\mathrm{term})(\Gamma'=\mathsf{cons}(\Gamma,x,A))
  \\{}& (d_1:\mathrm{deriv}(\Gamma,t,B)) \To
  (d_2:\mathrm{deriv}(\Gamma',t,B))
\end{align*}

\begin{example}
  From the empty repository, we build a derivation for $\lambda^{int} x. x$
  by constructing the repository:
  \begin{align}
    \Gamma &= \mathsf{nil} \\
    A &= \mathsf{int} \\
    x &= \mathsf{string}(\mathtt{"x"}) \\
    t_0 &= \mathsf{var}(x) \\
    \Gamma' &= \mathsf{cons}(\Gamma,x,A) \\
    V_0 &= \mathsf{zero}(\Gamma,x,A,\Gamma') \\
    D_0 &= \mathsf{init}(\Gamma,x,t_0,A,V_0) \\
    (t_1,C,D_1) &= \mathsf{typ\_lam}(\Gamma,x,A,A,t_0,\Gamma',D_0)
  \end{align}
  The last variable, $D_1$ is of type $\mathrm{deriv}(\Gamma,t_1,C)$.
  By unfolding the definitions and interpreting back the transformers,
  this expression becomes $\vdash\lambda^{int}x.x : int\to int$ which
  is what is expected.

\end{example}

In this example, we didn't use anything more than the transformers
already provided by the language construction, so this was just a
simple derivation. By now it should be clear that we can already build
whatever derivation we need: we just reified the object language into
the patch language by means of its constructors.

\begin{example}
  \label{ex:transf}
  Suppose now that we want to transform our program $D_1$ into a
  program $\lambda x^{int} f^{int\to int}.\ f x$. We need to transform
  the derivations as well. We add the following to our previous
  repository:
  \begin{align}
    f &= \mathsf{string}(\mathtt{"f"}) \\
    t_2 &= \mathsf{var}(f) \\
    \Gamma'' &= \mathsf{cons}(\Gamma',f,C) \\
    V_1 &= \mathsf{zero}(\Gamma',f,C,\Gamma'') \\
    D_2 &= \mathsf{init}(\Gamma'',f,t_2,C,V_1) \\
    D_3 &= \mathsf{weak}(\Gamma',f,C,A,t_0,\Gamma'',D_1) \\
    (t_3,D_4) &= \mathsf{typ\_app}(\Gamma'',t_2,t_0,A,A,C,D_2,D_3) \\
    (t_4,D_5) &= \mathsf{typ\_lam}(\Gamma',f,C,A,D_4) \\
    (t_5,D_6) &= \mathsf{typ\_lam}(\Gamma,x,A,D,D_5)
  \end{align}

  The resulting objects $t_5$ and $D_6$ are respectively interpreted
  back as the term $\lambda x^{int} f^{int\to int}.fx$ and its
  derivation $\vdash \lambda x^{int} f^{int\to int}.fx : int\to
  (int\to int)\to int$.
\end{example}

\subsection{Directions}

This formalization of proof/program transformations is a simple, yet
powerful idea, which opens numerous perspectives. The remaining
points of this sections will be directions for further investigations,
not definitive results \emph{per se}.

\paragraph{A basis for transformers}

In order to be able to represent \emph{any} transformation on the
object language $\mathcal L$, we need to specify the set of
metatheorems that we choose to promote as transformers. Let's try to
identify a minimal set of transformers able to represent all
transformations, a \emph{basis} so to speak for the construction of
the patch language.

One approach would be to identify, with the help of justified
examples, a set of common transformations on $\mathcal L$, as they are
used in practice when programming in $\mathcal L$. We would then need
to prove that they are \emph{complete} w.r.t all possible
transformations (more on completeness later). For a common, functional
programming language, we can imagine the following transformations:
\begin{itemize}
\item Rename a function or an argument, and all its subsequent
  occurrences in their scope
\item Add an argument to a function, and add a default parameter to
  all its calls
\item Replace a name by another name of the same type
\item add a constructor to a type declaration, adding a branch to all
  its \texttt{match\ldots with} constructs
\end{itemize}

We prefer a more low-level, syntactical approach based on the tree
representation of all objects, both syntactical and of typing: we
promote as transformers all operations of \emph{construction} and
\emph{destruction} on these trees, namely the constructors and
destructors of all syntactical objects, and the typing rules and their
inversion. Both these categories of operations manipulate directly the
tree structures (insert/delete a node) and provide a unifying way to
describe transformations. This way, we are not only guaranteed to be
complete, but also to help the design of an important piece of the
architecture of our system that what we will introduce later: the diff
algorithm.

\paragraph{Maximal sharing and completeness}

Variables assignments by transformer application is a simple way to
serve two different purposes. The first is to implement a dependent,
multi-hypothesis, multi-conclusion logic if we see transformers as
logical connectives. The second is to name all intermediate
construction steps of all objects, to allow free non-linear reuse of
previously constructed object so as to maximize sharing between
objects in a repository. This is the \emph{memoizing} function of
variable assignments, and it is the way we do incremental typing.

Maximal sharing of constructed object in a repository is a wanted
property. Not only does it ensures that we save the most work, which
is simply an efficiency concern, but it also ensures that we can
address objects by their variables: It states

\begin{definition}[Maximal sharing]
  A patch $\Delta$ has \emph{maximal sharing} if every object is
  assigned to a patch variable only once, namely that for all
  $x,y\in\Delta$, $\interp x_\Delta\neq\interp y_\Delta$.
\end{definition}

Knowing this, what is completeness in our system? A first, trivially
true and not very interesting statement is the following:
\begin{theorem}[Dummy completeness]
  All valid transformation from $\Delta$ to a program $p$ and its
  derivation $\vdash p : A$ are expressible as a program in $\mathcal
  P(\mathcal L)$.
\end{theorem}
\begin{proof}
  Since our patch language contains all typing rules, we just have to
  reconstruct the derivation of $\vdash p:A$ entirely (without using
  $\Delta$ in any ways).
\end{proof}

Obviously, this notion of completeness is not the one intended, as we
want to maximally reuse previous results from $\Delta$. It suggests
another notion of completeness, that we still to formalize and prove:

\begin{conjecture}[Completeness]
  All valid transformation from $\Delta$ to a program $p$ and its
  derivation $\vdash p : A$ are expressible as a program in $\mathcal
  P(\mathcal L)$ with maximal sharing.
\end{conjecture}

\paragraph{Bootstrapping the system}

We thus have a method, which can be automated, to generate from an
object language $\mathcal L$ and its typing rules, a patch language
$\mathcal P(\mathcal L)$ that, we saw, is also typed. Then nothing
prevents us from reapplying the functor $\mathcal P$ to the resulting
object. What do we obtain? Following what we just said, $\mathcal
P(\mathcal P(\mathcal L))$ is a language describing repository
transformations. For example, it should be able to describe the
application of patches at the end of a repository, the way we
implicitly did in our example \ref{ex:transf}.

Can this transformation help us describe what would be the core of a
version control system? To answer this, we would need to describe all
primitive operation of these systems (application, merge, commutation
of patches\ldots), and encode them as programs into the language. In
particular, we should try carefully to define a certain notion of
concurrency: what are patches that can be applied concurrently, what
are those that need a special sequentiality? Many interesting, open
questions concerning this bootstrapping arise.  Here are some of them:

\begin{itemize}
\item What metatheorem can we prove about the language $\mathcal
  P(\mathcal L)$
\item Which of them shall we choose as transformers for $\mathcal
  P(\mathcal P(\mathcal L))$?
\item Do they depend on the object language $\mathcal L$?
\item What can we say about the ``free'' language $\forall \mathcal X,
  \mathcal P(\mathcal X)$
\end{itemize}

\paragraph{A \emph{diff} algorithm}

The user of such a system surely does not want to write by hand what
we wrote in our previous example. It is both very verbose, as it shows
and names all intermediate constructs, and not useful as the
construction of a typing derivation is usually not his responsibility
but the work of the type inference algorithm. Moreover, one does not
want to change as drastically their habits as not to write
programs/proofs anymore, but only patches applied to previous versions
of its project.

Therefore, an interesting line of work would be to devise an algorithm
for computing the \emph{diff} between a repository and a file. It
would take a program and return the patch from the current repository
to that program. This patch could be typed afterwards, to guarantee
that the program given was well-typed with minimum effort.

If we see the patch language as a dependent, multi-hypothesis,
multi-conclusion logic, the work of searching for a resembles
proof-search as in \autoref{sec:auto}. Indeed, starting from a repository
$\Delta$ and having written the target program $t : A$, we are looking
for a composition of transformers $\Theta$ (viewed as logical
constants) realizing: \[ \Theta\quad:\quad\Delta \To (x = t) (d =
\mathrm{deriv(\Gamma_0,t,A)}) \] (the actual logic in which this proof
search is conducted still need to be specified formally).

But we have to be careful in this process, not to redo previous
searches, i.e. reusing maximally the objects of $\Delta$. This is
usually absent from traditional proof search, in which the proof
accounts only for its existence. Recall that we chose as transformers
the exact set of insert/delete operation on the derivation tree built.
Our process in this sense resembles much more a \emph{tree edit
  distance} algorithm: the best ``proof'' we can find is the one
minimizing the distance between (one of) the trees already built in
$\Delta$, and the new program we are willing to commit to $\Delta$. Of
course, these techniques will have to be adapted to the case of trees
with binders.

This algorithmic search is a challenging direction as it needs to be
very efficient, to be transparent to the user and not make us loose
the time we gained by typing incrementally. Moreover, the dual view of
this process as algorithmic content of proofs in automated proof
search seems a novel approach.

\paragraph{Functors as first-class patches}

Describing transformation of proofs/programs seems not only
interesting to implement a type-safe version control system as we
hinted above, but can be useful to encode other mechanisms commonly
used in proof development and programming, like modules.

Modules were designed as a tool for generic programming, and was
integrated in many functional programming languages, its archetype
being \textsf{SML}. One implements a given \emph{module} (a set of
functions and values), and give it a \emph{signature}, i.e. the bare
specification of its exported objects. Then, \textsf{functors} are
second-class functions, mapping a module of a certain signature to
another module. These functors can be applied, resulting in a new
module.

Given a language including modules but not functors, we conjecture
that we can represent functors as patches in the patch language.
Applying a functor would be the operation of applying the patch in a
given place in the repository.

Consider the following, toy language for libraries:
\begin{mathpar}
  A ::= \alpha\ |\ A\to A \and
  t ::= x\ |\ \lambda x. t\ |\ t\ t \\
  D ::= \{x:A; D\}\ |\ \cdot \and
  d :: \textsf{let}\ x=t; d\ |\ \cdot \\

  \infer[Nil]{ }{\Gamma\vdash\cdot\Rightarrow\cdot} \and
  \infer[Cons]{\Gamma\vdash t:A \\ \Gamma, x:A\vdash d\Rightarrow D}
  {\Gamma\vdash \textsf{let}\ x=t; d \Rightarrow\{x:A; D\}} \\
  \infer[Init]{(x:A)\in\Gamma}{\Gamma\vdash x:A} \and
  \infer[Lambda]{\Gamma;x,A\vdash t:B}{\Gamma\vdash\lambda x.t:A\to B}
  \and \infer[App]{\Gamma\vdash t:A\to B\\\Gamma\vdash
    u:A}{\Gamma\vdash t u:B} \and
\end{mathpar}

As opposed to the STLC, this language does not represents programs but
\emph{libraries} of programs (simple modules), having an
implementation ($\mathsf{let}\ x=t;\ \mathsf{let}\ y=u;\ \ldots$) and
a signature ($\{x:A;\ y:B;\ \ldots\ \}$).

A functor in this language can be encoded as a patch, taking as input
(i.e. having as free variable) a declaration $d$ of type the signature
$D$ expected, from which it constructs a typing environment $\Gamma$
and a new library, well-typed in $\Gamma$. This way, we encoded
functors as transformation operations on modules and we provided a
unified approach to them. Although this approach would require to be
formaized more carefully, we can already raise an interesting question
besides it: in the second-order language ($\mathcal P(\mathcal L)$) we
encoded functors, but we know that we can iterate this process
\emph{ad infinitum}. Can we then represent higher-order functors, like
functors taking functors as arguments, in $\mathcal P(\ldots (\mathcal
P(\mathcal L))\ldots)$?

\paragraph{User interaction}

Along with being a theoretical study on the impact of transformations
in proofs and programs, we believe that some practical application
could already come out, namely a simple, linear version control
system. We can already imagine the kind of user interaction this tool
would allow for the editing of developments: first of all, liberated
from the separate compilation paradigm, we don't edit whole files
anymore, but functions and values. For example, a user could ask:
\begin{verbatim}
> edit Set.add
\end{verbatim}
which would spawn a text editor with this function displayed (the
actual text being generated by pretty-printing the associated
\emph{term} sub-object of the current program). He then edit the
function and save the file. At this point, a patch is generated by the
\emph{diff} algorithm, possibly not well-typed. Then the patch is
temporarily committed (added to the repository at the current point),
and its type-checking begins. If it type-checks correctly, the patch
is definitively added and the system waits for a new user interaction.
If it doesn't type-checks, i.e. if there is a type error either in the
edited function or in the subsequent portion, then the errors are
displayed (possibly with additional context), and the user is asked
how to solve them. This is done either by hand, or by calling an
appropriate automatic strategy, defined by a high-level
transformer. Examples of high-level automatic resolution strategies
are:
\begin{description}
\item[$\alpha$-conversion] If we changed the name of an element of the
  context, we can decide to automatically rename all its occurrences;
\item[default argument] If we added an argument to a function, we can
  add a default parameter to all its occurrences;
\item[weakening] If the environment has been enlarged at some point,
  weakening allows us to modify any subsequent type derivation with
  the new environment;
\item[strengthening] If a name is not used in any part of the program,
  we can safely delete it from the environment;
\end{description}

\vspace{1em}

On the long term, we see other directions on this project. They
include:
\begin{itemize}
\item The use of special representations of the terms of the language
  using forward-pointers from binders to their occurrences,
\item The definition \emph{in the language} of patches of the
  higher-level transformers we described above,
\item Investigate the relationship with the version control tool
  \texttt{git}, as the structure of our repository resembles the
  \emph{object database} and its content-addressable namespace;
  possibly import ideas from it,
\item Approximation of type systems: when complex type systems are at
  stake (like dependent type theory), we could possibly devise a
  conservative approximation and use the real typing algorithm as a
  black box for retyping parts of the constructions, and still remain
  more efficient than typing the whole program,
\item Finally, we focused here on the semantic property of typing, but
  we could experiment other properties, like the preservation of
  operational semantics. The applications are numerous for programming
  (certified refactoring for example) but it seems more difficult
  because the base property is in the general case undecidable.
\end{itemize}

\bibliographystyle{plainnat}

\bibliography{../biblio}

% \appendix

% \section{ATP techniques}

% \subsubsection{Resolution}

% Resolution is the one, if not the most popular complete method for
% semi-deciding the refutation proof problem: given a statement, we can
% apply this method; if the statement is unsatisfiable in FOL, the
% method will eventually terminate with a refutation of it. The design
% and implementation of such a method is discussed at length in
% \cite{riazanov2002design}.  Many recent provers rely on resolution,
% which has proved empirically to be the most efficient.

% It was discovered in 1965 by \cite{robinson1965machine}, but since
% then many complete refinements have been proposed since its
% introduction: ordered resolution, selection\ldots (see
% \cite{bachmair2001resolution} for an overview). It consists of two
% phases: the transformation of the problem into clausal form, and the
% resolution itself, which is the iterated application of a (single,
% complete) rule mimicking the \emph{cut} rule of natural deduction.
% This is done with the help of a \emph{unification} algorithm to select
% relevant clause.

% Consider a set of formulas $H_i$ (the hypotheses) and a particular
% formula $G$ (the goal) of classical FOL. The satisfiability problem is
% the problem of finding assignations to the variables in $H_i, G$ such
% that the interpretation, in a boolean model satisfying all $H_i$, of
% the formula $G$ has truth value ``true''. Given a model of such type,
% it is easy to verify the satisfiability of the formula $G$ but the
% search for such a model is a very impractical exercise for a computer.
% We thus have to rely on \emph{syntactical methods} rather than such a
% semantic one, and resolution is one of them. It is easy to prove (by
% the De Morgan laws) that for all formula $B$, there is an
% equisatisfiable formula $B'$ which is in clausal normal form, i.e. has
% the form: \[ B = \bigwedge_{i=0}^{m}\left(\bigvee_{j=0}^n \pm
%   A_{ij}\right) \] where $A_{ij}$ are all \emph{literals}, i.e. terms
% of the signature and all variables in the literals are implicitly
% universally quantified.  Moreover, the satisfiability of the goal $G$
% provided that all $H_i$ are satisfiable is equivalent to the
% satisfiability of: $$\bigwedge_i(-H_i) \vee G$$ This allows us to work
% directly on sets of \emph{clauses}, or disjunctive set of literals and
% treat the literal as plain terms, without having to consider their
% semantics.

% Now let's take the negation of our starting formulas, and put it into
% clausal normal form. The resolution procedure works by iteratively
% enlarging this set of clauses with new clauses (the resolvants) until
% it generates an empty clause. The empty clause, writted $\bot$,
% denotes the unsatisifability of the clause, thus the unsatisifability
% of the whole set of clause, and if the method is sound, deriving
% $\bot$ means that the starting set of clause was unsatisfiable, i.e.
% that our formula is satisfiable, since the negation of a satisfiable
% formula is unsatisfiable. Here are the two rules of resolution:

% \begin{mathpar}
%   \infer[Resolution]{C,+A \and C',-A'}{(C, C')\sigma}\ 
%   \sigma=mgu(A,A') \and \infer[Factoring]{C,+A,+A'}{(C,+A)\sigma}
%   \ \sigma=mgu(A,A')
% \end{mathpar}

% The function $mgu$ computes the \emph{most general unifier} of two
% literals, and the resolution rule states that two unifiable literals
% of inverse signs in two sets of clauses can help generate a new clause
% formed by the concatenation of the previous sets, provided that we
% instantiate their variable by the result of the unification. The
% factoring rule, which is unnecessary in the degraded case of Horn
% clauses (clause with \emph{at most} one positive literal) but
% necessary for general clauses, states that two positive (resp.
% negative) unifiable literals can be factorized into one, provided that
% we instantiate the rest of the clause. We see that only a resolution
% step can lead to the final result: in the case where both premises
% are singleton clauses containing two unifiable literals, resolution
% leads to the empty clause.

% This method enjoys soundness and completeness.

% \begin{theorem}[Soundness]
%   Let $S$ be a set of clauses and $S'$ a set obtained by resolution from
%   $S$. All models of $S$ are models of $S'$.

%   In particular, if $\bot$ is deducible from $S$, then $S$ is
%   unsatisfiable.
% \end{theorem}

% Soundness is trivial to prove: both rules are interpreted valid in any
% model of first-order logic. From a proof-theoretic point-of-view, the
% resolution rule is much alike to the \emph{cut} rule of sequent
% calculus, and the factoring the contraction rule:
% \begin{mathpar}
%   \infer[Cut]{\Gamma,A \vdash \Delta \and \Gamma\vdash A,\Delta}
%   {\Gamma\vdash\Delta} \and
%   \infer[CR]{\Gamma\vdash\Delta, A, A}{\Gamma\vdash\Delta, A}
% \end{mathpar}

% \begin{theorem}[Completeness]
%   Let $S$ be a set of clause. It is unsatisfiable \emph{iff} we can
%   derive $\bot$ by the only application of \textsc{Resolution} and
%   \textsc{Factoring}.
% \end{theorem}

% Completeness is harder and requires complex proof techniques. A
% popular one is the \emph{semantic tree method}: we construct a
% \emph{Herbrand tree} of the set of clause, and consider the
% \emph{failure} nodes where the set becomes unsatisfiable. We see that
% resolution let these nodes climb up the tree, until it arrives to the
% root. Thus we prove that the iterated application of resolution is
% terminating, and that for an unsatisfiable $S$, $\bot$ is eventually
% derived. Note that this does not guarantee the termination of the
% process in all cases: if the given clauses are satisfiable, it
% \emph{will} loop forever. See \cite{goubault2005blossom} for an
% overview of this proof method.
% \\[-0.5em]

% The high non-determinism of this rule makes the bare resolution rather
% inefficient. Indeed, the number of clause in the system during
% resolution keeps growing exponentially: each application of resolution
% of factoring leads to a new clause which is added in the global pool
% and ready to be in turn combined with other at the next iteration. The
% search space is huge (unless a trivial case is met), and refinements of
% the resolution principle have been devised to reduce it when possible,
% i.e. by still remaining complete.

% \paragraph{Ordered resolution} One of the inefficiencies of this
% method is the possibility to combine every literal of a clause with
% every literal from another (or possibly the same) clause. It is
% possible to restrict the application of resolution to \emph{maximal}
% literal in every clause, according to an ordering on terms respecting
% some criteria. The rule becomes:
% \[ \infer {C,+A_1,\ldots,+A_n \and C', -A'}
% {(C,C')\sigma}\quad\sigma=mgu(A_1,\ldots,A_n,A') \] where $n\geq 1$,
% and $A_1\ldots A_n$ are all maximal in the first premise, $A$ in the
% second with respect to a strict, stable order $\succ$. Stable means
% that $A\succ B$ implies $A\sigma\succ B\sigma$ for all $A,B$ literals
% and $\sigma$ substitution. We emit the same restriction for the
% factoring rule.

% In practice, stable orders are chosen among reduction orderings, i.e.
% ordering expressing the fact that a term is ``smaller'' than another,
% so that unification has to do the minimum job possible.

% \paragraph{Ordered resolution with selection} An even bigger
% restriction on the application of the resolution principle is to
% parameterize it with respect to a special \emph{selection function}
% $sel : C \to \wp(C)$ that selects negative literals, whose role is to
% select a default choice of literals to resolve in a clause. The rule
% becomes:
% \[ \infer
% {\{C_i,A_{i1},\ldots,A_{in_i}\}_{1\leq i\leq l} \and
%   C',-A'_1,\ldots,-A'_l}
% {(C_1,\ldots,C_l,C')\sigma}
% \]
% where:
% \begin{enumerate}
% \item $n_i\geq 1$ for all $i$, $1\leq i\leq l$,
% \item $\sigma = mgu\{ A_{ij} = A'_i\}_{1\leq i\leq l,\ 1\leq j\leq n_i}$,
% \item $sel(C_i, A_{i1},\ldots,A_{in_i}) = \emptyset$ and
%   $A_{i1},\ldots,A_{in_i}$ maximals in their clause,
% \item either $sel(C',-A'_1,\ldots,-A'_l) = -A'_1,\ldots,-A'_l$ with
%   $l\geq 1$, or $sel(C',-A'_1,\ldots,-A'_l) = \emptyset$ and $-A'1$
%   maximal in its clause.
% \end{enumerate}

% As counter-intuitive as it may sound, even with an arbitrary selection
% function, this method remains complete.
% \\[-0.5em]

% As we have seen, these method of proof search are based on the
% decomposition of a first-order formula into a set of clauses by
% applying the De Morgan laws. Most of these laws are well-known to be
% false in intuitionistic logic, in fact most of them are true only in
% one direction. One could easily drop to the conclusion that these
% methods are inapplicable to an intuitionistic framework. In fact, it
% turns out that the process of resolution in itself is well-applicable
% in this case, and the preprocessing into clausal normal form is only a
% commodity. One could translate this process into a fully
% intuitionistic one by basing the system, not on clausal forms anymore,
% but on implicative form. What we get by doing this is a system that
% applies its hypothesis to the goal creating new goals, until all of
% them vanish. We will discuss this approach in more depth in
% \ref{sec:psitp}.

% \subsubsection{Equality treatment}

% Equality is an ubiquitous notion in mathematics. Yet, FOL doesn't
% include any special treatment for this symbol: it is defined
% axiomatically inside the logic (as the smallest congruent equivalence
% relation), which makes it very inefficient to use in practice: proving
% an equality triggers the blind application of the transitivity and
% symmetry lemmas in all possible ways:
% \begin{align*}
%   x=x & \qquad (reflexivity)\\
%   x=y \to y=x & \qquad (symmetry) \\
%   x=y \to y=z \to x=z & \qquad (transitivity)\\
%   x=y \to f(x)=f(y) & \qquad (congruence,\ \forall f)
% \end{align*}

% A better way to treat the equality is by means of rewriting. It is the
% goal of paramodulation, superposition and their derivatives. They all
% are inference rules for treating equality externally to the logic, as
% a mean to replace equals by equals in formulas. Again, a wealth of
% variants have been devised from the main, relatively inefficient idea:
% E-unification (paramodulation treated inside the unification
% algorithm), oriented paramodulation with selection. Completion is an
% important ingredient to the treatment of equality: Given a set of
% equalities, it is an algorithm for transforming it, when possible,
% into a set of confluent rewrite rules.  This accelerates the procedure
% of paramodulation without loss of completeness: we know that for some
% rules rewriting isn't necessary in both directions. Finally,
% congruence closure is a decision method for equalities in ground
% equational theories, used in practice in the SMT provers, a new
% generation of provers that has proved to be very efficient.

% \paragraph{Paramodulation}

% Paramodulation (See \cite{nieuwenhuis2001paramodulation} for a
% complete survey) originated as a development of resolution, introduced
% soon after it by \cite{wos1968paramodulation}: for improving
% resolution-based methods, the study of the equality predicate has been
% particularly important because its axiomatic treatment was unusable.
% This lead to the design of an additional, dedicated inference rule:
% \[ \infer{C, s=t \and D}{(C, D[t]_p)\sigma}\quad\sigma=mgu(s,D|_p)\]
% where $D|_p$ is the subterm at position $p$, and $D[t]_p$ denotes the
% result of replacing in $D$ the subterm at position $p$ with $t$. It is
% based on Leibnitz's law of replacing equal by equal: every instance of
% a member of an equation inside a clause can be rewritten to the other
% member of the equation. This applies as well to clauses containing
% equations. The addition of this rule to general resolution was proven
% to be a complete way to semi-decide FOL with equality, but, the rule
% being so general, it is almost unusable as-is.

% Later on, it was proved that paramodulation was unnecessary to perform
% into variables, that is, paramodulations where $D|_p$ is a variable.
% However, even under these restrictions, paramodulation is difficult to
% control: unless additional refinements are considered, it quickly
% produces a large amount of unnecessary clauses, expanding the search
% space excessively.

% \paragraph{Superposition}

% Superposition is a way to control the rule of paramodulation by
% restricting the application of the rule with respect to an
% \emph{order}, similarly to ordered resolution. The intuition behind
% superposition is to only rewrite \emph{big} terms into \emph{smaller}
% ones, these predicates being defined by a special \emph{rewriting
%   order} (a monotone, stable, well-founded order). Superposition
% consist in these two rules, parameterized on the order $\succ$:
% \begin{align}
%   \infer{C, l=r\and C',\pm (s=t)} {C,C', \pm (s[r]_p=t)}\qquad
%   &\text{if}\quad
%   \begin{array}{l}
%     s|_p\cong l, l\succ r, s\succ t \\
%     l\succ u\text{ for all } u\in C \\
%     s\succ v\text{ for all } v\in C' 
%   \end{array} \\
%   \infer{C, - (s = s)}{C}\qquad
%   &\text{if}\quad\ \ 
%   s\succ v \text{ for all } v\in C
% \end{align}

% Numerous term orderings have been proposed in the literature that
% tackle the problem of finding an ordering compatible with the
% requirements of superposition, general enough to capture the intuitive
% notion of the size of a term, and efficiently computable. Among the
% \emph{syntactic} term orderings, the Recursive Path Ordering (RPO) and
% the Knuth-Bendix ordering are the most successful. They both contain
% the subterm order, and satisfy the required properties: they are
% congruences (or monotone), stable under substitution and
% well-founded. By their syntactical nature, they are easily computable.
% \emph{Semantic} term ordering are rewriting orders based on semantical
% properties rather than their syntactic form. It usually requires to
% find a model of the pair of term which is often undecidable but can be
% approximated. Examples of such models are graph models (dependency
% pairs) or polynomial models (which requires to solve diophantine
% equations).

% \paragraph{Completion}

% The first use of the superposition rule has been by Knuth and Bendix
% for \emph{completion}. Completion is a procedure that attempts to
% transform a given rewrite system into an equivalent confluent one. It
% finds critical pairs in the rewrite system by superposing its
% equations, orienting them according to the given order. During the
% completion procedure, equations are simplified by rewriting and
% tautologies are removed. For example, group theory can be defined as
% the set of equations:
% \[ E\ :=\ \{ x\cdot e=x; I(x)=e; x\cdot(y\cdot z)=(x\cdot y)\cdot z
% \} \] and is transformed by the completion procedure into the
% confluent rewrite system:
% \[ R\ :=\ \{ x\cdot e\to x; I(x)\to e; x\cdot(y\cdot z)\to (x\cdot
% y)\cdot z \} \]

\end{document}