### 2011-01-14 papers/ICFP2011/paper.tex
ICFP: C'est parti!
git-svn-id: http://gasp.dev.regis-gianas.com/svn/trunk@234 403e655e-980a-47bd-8c60-1ee5348bb6b7

\documentclass[preprint]{sigplanconf}

\usepackage{amsmath}
\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{color}
\usepackage[pdftex,backref=page,colorlinks=true]{hyperref}

\definecolor{bwgreen}{rgb}{0.183,1,0.5}
\definecolor{bwmagenta}{rgb}{1,0.169,0.909}
\definecolor{bwblue}{rgb}{0.317,0.161,1}
\hypersetup{
  linkcolor=blue,
  citecolor=blue
}

\newcommand{\remplan}[1]{\noindent\textcolor{bwblue}{$\triangleright$ \textbf{#1}}}
\newcommand{\remtext}[1]{\textcolor{bwgreen}{$\triangleright$ \textsl{#1}}}

\begin{document}

\conferenceinfo{ICFP '11}{June 4--8, Tokyo, Japan} 
\copyrightyear{2010}
\copyrightdata{[to be supplied]}

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{TBD}
%\subtitle{Subtitle Text, if any}

\authorinfo{Matthias Puech}
           {University of Bologna, University Paris Diderot}
           {puech@cs.unibo.it}
\authorinfo{Yann Régis-Gianas}
           {University Paris Diderot, CNRS, and INRIA}
           {yrg@pps.jussieu.fr}

\maketitle

\begin{abstract}
This is the text of the abstract.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

\terms
term1, term2

\keywords
keyword1, keyword2

\section{Introduction}

\remtext{Une intro que j'avais écrit il y a longtemps et que je sais
  pas où mettre. Sûrement à jeter\ldots}

Programs are nowadays faster to write and safer to execute, thanks to
high-level programming concepts and expressive type systems. Still,
one aspect of programming didn't change very much---if not at
all---since the early days of high-level programming: we still write
programs in text files and when we are done, we let the compiler check
and compile them.  Though, every programmer knows that writing a
program longer than a few dozen of lines is a highly non-linear task
involving constant experiments, fixes, rollback on previous
modifications etc., all of those eventually validated by some a
posteriori criterion: execution and ``empirical'' test, and
increasingly type-checking. In a nutshell, programmers spend more time
\emph{editing} than \emph{writing}. Yet, this aspect of their
day-to-day workflow is not, or poorly taken into account in a
developer's toolchain:
\begin{itemize}
\item To avoid recompiling a whole program upon a change, we split
  them into files, and (re-)compile them separatly. Very often yet,
  this separation serve different purpose: abstraction, splitting into
  logical units, namespace management\ldots\ Why should these purposes
  coincide?
\item To keep track of past versions of a program, we use text-based
  version control systems that have no awareness of the actual content
  of the file. In a sense, the non-linearity of the task is managed
  outside the scope of the language, by manual, textual
  transformations. Isn't there a lot to gain in making these tools
  syntax-aware, even semantics-aware?
\item As a result, all the semantic information that could be
  available to the programmer and act as a \emph{guidance tool} to
  write programs (e.g. type informations) are only used by the
  type-checker as a \emph{repressive} measure against programming
  errors (type errors).
\end{itemize}

These observations apply to the usual, wide-spread programming
languages, but even if the current toolchain in use up to now was a
reasonable approximation to the incremental interaction with the
compiler intended by the user, they become \emph{a fortiori} necessary
when it come to the new, demanding languages embedding formal
verification aspects, like proof assistants or rich specification
languages. In these emerging languages, interaction with a type- or
proof-checker is unavoidable due to the difficulty of writing correct
proofs without guidance from the system, and the constant modification
of the source makes necessary to observe incrementally the effect of a
small change on the whole edifice. Moreover, as the complexity of the
object to check increases, we cannot afford anymore to rely on the
usual unit of compilation, the file, a need a finer-grained tool.

We propose to devise here a data structure for repositories of
programs and proofs allowing to take advantage of the incremental
nature of a programmer's workflow.

% The problem (1 page)
% My idea (2 pages)
% The details (5 pages)
% related works (1-2 pages)
%conclusion and further work (0.5 page)

\bibliographystyle{abbrvnat}

\end{document}