### 2011-10-07 doc/papers/TLDI2012/abstract.tex
started TLDI abstract

\documentclass[9pt,authoryear]{sigplanconf}

\usepackage{ntheorem}
\usepackage{amsmath}
\usepackage{amssymb}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\usepackage{mathpartir}
\usepackage{macros}
\usepackage{breqn}

\begin{document}

\conferenceinfo{TLDI '12}{January 28 2012, Philadelphia, USA.}
\copyrightyear{2012} 
\copyrightdata{[to be supplied]}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Incremental Type Checking}
%\subtitle{Subtitle Text, if any}

\authorinfo{Name1}
           {Affiliation1}
           {Email1}
\authorinfo{Name2\and Name3}
           {Affiliation2/3}
           {Email2/3}

\maketitle

\begin{abstract}
  We study the problem of verifying the well-typing of terms, not in a
  batch fashion, as it is usually the case for typed languages, but
  incrementally, that is by sequentially modifying a term, and
  re-verifying each time only a smaller amount of information.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

\terms
term1, term2

\keywords
keyword1, keyword2

\section{Introduction}

As developments grow and type systems become more involved, writing a
correct program in one shot becomes quite difficult. On the other
side, writing a program in many correct steps can be tedious because
generally, the verification tools rechecks the whole development at
each step. This second option is the usual practice when the time for
verification is neglectible wrt. the time for making the change, but
this is less and less the case, especially when the language in
question embed proof aspects, and verification involves proof
search. Some mechanisms exist already to cope with the incrementality
of proofs or program development: separate compilation, interactive
toplevel with undo, tactic languages; they all provide in different
ways a rough approximation of the process of modifying and checking
incrementally a large term.

We propose here an architecture for a generic incremental
type-checker, a data structure for repositories of typed proofs and a
language for describing proof deltas. It is based on the simple idea
of sharing common subterms to avoid recheckings and exploits the idea
of encoding a derivation in a metalanguage. This way, given a
signature declaring the typing rules and an (untrusted) typing
algorithm for my language of choice, I get an incremental type-checker
for that language. The metalanguage approach gives us the ability to
encode all the aforementioned usual incrementality mechanisms, and
more.

\section{Building up the metalanguage}

As a first example, let us consider a purposedly simplistic sorted
language of boolean and arithmetic expressions:
$$ e, e' \gequal n \gor e + e' \gor e \land e' \gor e \leq e' $$
The algorithm to determine in a batch fashion wether the term
$$ e_1 = (1 + 3 \leq 2 + 4) \land (8 \leq 3) $$
is well-sorted is trivial (we don't care about its evaluation here,
just well-sortedness). But what if I change subterm $2+4$ in $e_1$
into $7 \leq 2+4$, to obtain $e_2$? Clearly, it should be verified
that context $7\leq []$ is well-sorted (it is), that $2+4$ ``fits''
into its hole (it does), that the whole thing ``fits'' into its new
context $(1+3\leq [])\land(8\leq 3)$ (it is not); but the other,
unchanged subterms need not be. For that, we would have to
``remember'' the state of the verifier in some way.

If only we had \emph{names} (memory addresses, hashes) for
sufficiently enough subterms of our initial term, $$ e_1 =
(\overbrace{1 + 3}^\mmeta \leq \overbrace{2+4}^\mmmeta) \land
\overbrace{(8 \leq 3)}^\mmmmeta\text{\ ,} $$ we could express
concisely the change as a \emph{delta} $$\delta_1 = (\mmeta\leq(7\leq
\mmmeta))\land \mmmmeta \text{\ ,}$$ sharing unchanged subterms. If
only we had \emph{annotated} our initial term with the states of the
verifier,
$$e_1 = (\overbrace{1 + 3}^{\mmeta:\nat} \leq \overbrace{2 + 4}^{\mmmeta:\nat}) \land
\overbrace{(8 \leq 3)}^{\mmmmeta:\bool}\text{\ ,} $$ we would have a
simple process to verify $e_2$ taking advantage of $e_1$'s derivation,
in $O(|\delta_2|)$: verify $\delta_1$ as a term, retrieving the sort
of names from a stored map. This suggests a data structure for a
\emph{repository} of named, annotated, verified subexpressions: a map
from names, or \emph{metavariables}, to terms and types, together with
a \emph{head} metavariable identifying the top of the term: $$\mr
\gequal X, \Delta \quad\text{where}\quad \Delta : (\mmeta\mapsto M :
A) \text{\ .}$$

Terms $M$ should be written in a language allowing to encode our expressions with metavariables,
and types $A$ should contain the whole state of the batch verifier
(here the sort). One modular choice for this metalanguage is \emph{LF}
\cite{harper1993framework}: it allows to specify syntax and rules of
an object language as a \emph{signature} $\Sigma$, and check terms
against this signature. We'll use an increasing fragment of it. The
so-called \emph{intrinsic} style of LF signature for our language
could be:
\begin{align*}
  tp &: \type,\quad nat : tp,\quad bool : tp,\quad exp : tp\to nat,\\
  atom &: \nat\to exp\ nat,\\
  plus &: exp\ nat \to exp\ nat \to exp\ nat,\\
  and &: exp\ bool\to exp\ bool\to exp\ bool,\\
  leq &: exp\ nat\to exp\ nat\to exp\ bool
\end{align*}
In this style, the encoding of an expression and its (object) type is
a term in the metalanguage, but appears in the type of this
expression. The repository associated with expression $e_1$ is
\begin{align*}
  T,
  \begin{array}{rl}
  X &\mapsto plus\ 1\ 3 : exp\ nat \\
  Y &\mapsto plus\ 2\ 4 : exp\ nat \\
  Z &\mapsto leq\ 8\ 3 : exp\ bool \\
  T &\mapsto and\ (leq\ X\ Y)\ Z : exp\ bool \\
\end{array}
\end{align*}

\section{Expressivity}

\section{Architecture}

The \emph{kernel} is the component in charge of verifying terms
against a signature and a repository, and updating this repository. It
supports two basic operations:
\begin{itemize}
\item $\pfunction{push}{\Sigma}{\mr,M}$ checks $M$ against $\Sigma$
  in $\mr$, synthetizes its type $A$, chooses a fresh metavariable
  $\mmeta$ for $M$ and returns $\mr[\mmeta\mapsto M:A]$ and $\mmeta$.
\item $\pfunction{pull}\Sigma{\mr, \mmeta}$ returns the term $M$
  associated with $\mmeta$ in $\mr$ recursively: all metavariables are
  unfolded to their definitions.
\end{itemize}

The \emph{slicer} is the component in charge of slicing a term $M$
into many terms for the kernel to push in the repository.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrvnat}
\bibliography{../../../../../Documents/These/biblio.bib}

% \begin{thebibliography}{}
% \softraggedright
% \bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
% P. Q. Smith, and X. Y. Jones. ...reference text...
% \end{thebibliography}

\end{document}