### 2011-10-10 doc/papers/TLDI2012/abstract.tex
TLDI: version finale

\documentclass[9pt]{sigplanconf}

\usepackage{ntheorem}
\usepackage{amsmath}
\usepackage{amssymb}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\usepackage{mathpartir}
\usepackage{macros}

\begin{document}

\conferenceinfo{TLDI '12}{January 28 2012, Philadelphia, USA.}
\copyrightyear{2012} 
\copyrightdata{[to be supplied]}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Safe Incremental Type Checking}
%\subtitle{Subtitle Text, if any}

\authorinfo{Matthias Puech}
           {Department of Comp. Sci., Univ. of Bologna,\\
             PPS, Team $\pi r^2$ (Univ. Paris Diderot, CNRS, INRIA)}
           {puech@cs.unibo.it}

\authorinfo{Yann R\'egis-Gianas}
           {PPS, Team $\pi r^2$ (Univ. Paris Diderot, CNRS, INRIA)}
           {yrg@pps.jussieu.fr}

\maketitle

\begin{abstract}
  We study the problem of verifying the well-typing of terms, not in a
  batch fashion, as it is usually the case for typed languages, but
  incrementally, that is by sequentially modifying a term, and
  re-verifying each time only a smaller amount of information than the
  whole term, still ensuring that it is well-typed.
\end{abstract}

\category{D.3.3}{Language Constructs and Features}{Data types and
  structures} \category{F.3.1}{Logics and Meanings of
  Programs}{Specifying and Verifying and Reasoning about
  Programs --- Logics of programs}

\terms
Theory, Languages

\keywords
incrementality, type checking, logical framework, version control

\section{Introduction}

As programs grow and type systems become more involved, writing a
correct program in one shot becomes quite difficult. On the other
hand, writing a program in many correct steps is the usual practice
when the time for verification is negligible; the verification tool
then rechecks the entire development at each step. But this gets more
tedious especially when the language in question contains proof
aspects, and verification involves proof search. Some mechanisms
already exist to cope with the incrementality of proofs or program
development: separate compilation, interactive toplevel with undo,
tactic languages; they all provide in different ways a rough
approximation of the process of modifying and checking incrementally a
large term.

We propose here an architecture for a generic and safe incremental
type checker, a data structure for repositories of typed proofs and a
language for describing proof deltas. It is based on the simple idea
of sharing common subterms to avoid rechecking, and exploits encoding
a derivation in a metalanguage to guarantee the well-typing of the
result. This way, given a signature declaring the typing rules and an
(untrusted) typing algorithm for my language of choice, I get an
incremental type checker for that language. The metalanguage approach
gives us the ability to encode all the aforementioned usual
incrementality mechanisms in a type-safe way, and more, making our
system akin to a typed version control system.

\section{Sharing-based incrementality}

As a first example, let us consider a purposedly simplistic sorted
language of boolean and arithmetic expressions:
$$ e, e' \gequal n \gor e + e' \gor e \land e' \gor e \leq e' $$
The algorithm to determine in a batch fashion whether the term
$$ e_1 = (1 + 3 \leq 2 + 4) \land (8 \leq 3) $$ is well-sorted is
trivial (we don't care about its evaluation here, just its
well-sortedness). But what if I then change subterm $2+4$ in $e_1$
into $7 \leq 2+4$, to obtain $e_2$? Clearly, it should be verified
that context $7\leq []$ is well-sorted (it is), that $2+4$ ``fits''
into its hole (it does), that the whole expression ``fits'' into its
new context $(1+3\leq [])\land(8\leq 3)$ (it does not); but the other,
unchanged subterms need not be verified again. To achieve this
incremental verification, the system would have to ``remember'' the
states of the verifier in some way.

If only we had \emph{names} (memory addresses, hashes) for enough
subterms of our initial term, $$ e_1 = (\overbrace{1 + 3}^\mmeta \leq
\overbrace{2+4}^\mmmeta) \land \overbrace{(8 \leq 3)}^\mmmmeta\text{\
  ,} $$ we could express concisely the change as a
\emph{delta} $$\delta_1 = (\mmeta\leq(7\leq \mmmeta))\land \mmmmeta
\text{\ ,}$$ using these names to refer to unchanged subterms. If only
we had \emph{annotated} our initial term with the states of the
verifier,
$$e_1 = (\overbrace{1 + 3}^{\mmeta:\nat} \leq \overbrace{2 +
  4}^{\mmmeta:\nat}) \land \overbrace{(8 \leq
  3)}^{\mmmmeta:\bool}\text{\ ,} $$ we would have a simple process to
verify $e_2$ taking advantage of $e_1$'s derivation, in
$O(|\delta_2|)$: verify $\delta_1$ as a term, retrieving the sort of
names from a stored map. This suggests a data structure for a
\emph{repository} of named, annotated, verified subexpressions: a
monotonously growing map, from names, or \emph{metavariables}, to terms
and types, together with a \emph{head} metavariable identifying the
top of the expression: $$\mr \gequal X, \Delta \quad\text{where}\quad
\Delta : (\mmeta\mapsto M : A) \text{\ .}$$

\section{A metalanguage to encode derivations}

What language are terms $M$ and types $A$ written into? Terms should
encode our expressions with metavariables, and types $A$ should encode
the whole state of the batch verifier (here the sort). An obvious
choice is to take $M\gequal(e \text{ with metavariables})$ and
$A\gequal \bool\gor\nat$, but switching to another language, we'd have
to redefine another repository language. Moreover, this choice is no
longer expressive enough when introducing binders. Another more
modular choice for this is the \emph{metalanguage} \emph{LF}
\cite{harper1993framework}: it allows to specify syntax and rules of
an object language as a \emph{signature} $\Sigma$, and check terms
against this signature with a generic algorithm. We'll use an
increasing fragment of it. The so-called \emph{intrinsic} style of LF
signature for our expression language is:
$$
\begin{array}{rl}
  \mathsf{tp} &: \type,\quad
  \mathsf{nat} : \mathsf{tp},\quad
  \mathsf{bool} : \mathsf{tp},\quad
  \mathsf{exp} : \mathsf{tp}\to \mathsf{nat},\\
  \mathsf{atom} &: \nat\to \mathsf{exp}\ \mathsf{nat},\\
  \mathsf{plus} &: \mathsf{exp}\ \mathsf{nat} \to \mathsf{exp}\ \mathsf{nat} \to \mathsf{exp}\ \mathsf{nat},\\
  \mathsf{and} &: \mathsf{exp}\ \mathsf{bool}\to \mathsf{exp}\ \mathsf{bool}\to \mathsf{exp}\ \mathsf{bool},\\
  \mathsf{leq} &: \mathsf{exp}\ \mathsf{nat}\to \mathsf{exp}\ \mathsf{nat}\to \mathsf{exp}\ \mathsf{bool}
\end{array}
$$
In this style, both the encoding of an expression and its sort are
\emph{terms} in the metalanguage, but the sort appears in the
\emph{type} of the encoded expression. As an example, the repository
associated with expression $e_1$ is
\begin{align*}
  T, \left(
  \begin{array}{l}
  X \mapsto \mathsf{plus}\ 1\ 3 : \mathsf{exp}\ \mathsf{nat} \\
  Y \mapsto \mathsf{plus}\ 2\ 4 : \mathsf{exp}\ \mathsf{nat} \\
  Z \mapsto \mathsf{leq}\ 8\ 3 : \mathsf{exp}\ \mathsf{bool} \\
  T \mapsto \mathsf{and}\ (\mathsf{leq}\ X\ Y)\ Z : \mathsf{exp}\ \mathsf{bool} \\
\end{array}\right)
\end{align*}

The dependent nature of types in LF allows to express more complex
languages. We can for example add functions, applications and
variables to our expressions in a purely first-order style (using de
Bruijn indices for variables) if we annotate them not only with sorts
but with an environment of free variables:
$$
\begin{array}{rl}
  \mathsf{exp} &: \mathsf{env}\to \mathsf{tp}\to \type,\\
  \mathsf{atom} &: \prd E {\mathsf{env}} \nat\to \mathsf{exp}\ E\ \mathsf{nat},\\
  \mathsf{var} &: \prd E {\mathsf{env}} \prd A {\mathsf{tp}} \mathsf{var}\ E\ A \to \mathsf{exp}\ E\ A,\\
  \mathsf{leq} &: \prd E {\mathsf{env}} \mathsf{exp}\ E\ \mathsf{nat}\to \mathsf{exp}\ E\ \mathsf{nat}\to \mathsf{exp}\ E\ \mathsf{bool}, \\
  \mathsf{lam} &: \prd E {\mathsf{env}} \prd {A, B} {\mathsf{tp}}
  \mathsf{exp}\ (\mathsf{cons}\ A\ E)\ B \\ &
  \qquad\to\mathsf{exp}\ E\ (\mathsf{arr}\ A\ B)\\
  \ldots
\end{array}
$$
The encoded expressions are however very verbose: each term
constructor takes as argument all these annotations. We can
nonetheless make these information \emph{implicit} in terms (but
explicit in types) and let a reconstruction algorithm infer them, as
in \cite{necula1997efficient}. This reconstruction is
language-dependent, user-provided but does not impair the safety of
the system for the whole term is still checked afterwards.

LF promotes the use of lambda-tree syntax to represent binders:
instead of encoding the syntax first-order, it uses the $\lambda$
binder built in LF to encode binders in the object language. This
style of encoding has the advantage of making the manipulation of the
environment (weakening, exchange\ldots) implicit in the deltas, but
raises new challenges for the delta language and the verification
process: how to share a subterm underneith a lambda? How to
efficiently verify that such a delta is well-typed?

\section{Expressivity}

Aside from enabling to encode a large class of deductive systems
safely and generically, the metalanguage approach allows to express
incrementality features usually implemented in an ad-hoc manner,
simply by adding new constants to the signature.

\paragraph{Version control}

Suppose we want to implement an \emph{undo system}, storing successive
versions of a closed expression of sort $bool$ and able to rollback to
a previous version. We add constants
$$
\begin{array}{rl}
  &\mathsf{version} : \type,\quad \mathsf{vnil} : \mathsf{version},\\
  &\mathsf{vcons} : \mathsf{exp}\ \mathsf{nil}\ \mathsf{bool} \to \mathsf{version} \to \mathsf{version}
\end{array}
$$
to the signature. The empty repository is now represented as
$vnil$. Each time we have pushed a full expression $M$, and if $S$ was
the previous head (a $\mathsf{version}$ called its \emph{ancestor}), we push
$\mathsf{vcons}\ M\ S$. This gives us a data structure for an undo stack, and a
\emph{commit} algorithm. But the sharing inherent to our repositories
lets us actually represent \emph{trees} of versions, by sharing common
stack tails, each list head being a \emph{branch}. Reconciling two
branches' changes into a unique head is called \emph{merging} in
version control system's terminology: a merge is a version with
several ancestors. We can represent merges by revising our previous
addition to the signature into
$$\begin{array}{rl}
  &\mathsf{version} : \type,\quad
  \mathsf{ancestors} : \type,\quad
  \mathsf{anil} : \mathsf{ancestors},\\
  &\mathsf{acons} : \mathsf{version} \to \mathsf{ancestors} \to \mathsf{ancestors},\\
  &\mathsf{vcons} : \mathsf{exp}\ \mathsf{nil}\ \mathsf{bool} \to \mathsf{ancestors} \to \mathsf{version}
\end{array}$$
This defines a data structure to represent (acyclic) \emph{graphs} of
versions; it is the exact data structure of repository used by version
control systems \textsf{Git}, \textsf{Monotone} and \textsf{Mercurial}
(see e.g. \cite{chacon2009git}) except that where they have directories
and text files we have arbitrary typed terms.

\paragraph{Top-down construction}

While our system is based on \emph{bottom-up} term construction, we
can encode \emph{top-down} construction common to some programming
environments (e.g. \textsf{Agda}) and tactic-based proof assistants
(e.g. \textsf{Coq}). The user constructs terms by
successively filling \emph{holes} with terms containing other
holes. To add (linear) holes to our expressions, add constant
$$ \mathsf{hole} : \prd E {\mathsf{env}} \prd A {\mathsf{tp}} \mathsf{exp}\ E\ A $$
to the signature. To instantiate a hole with an expression, commit the
substituted term preserving sharing of subterms.

\section{Architecture}

We can implement this system following a layered architecture.

The \emph{kernel} is the component in charge of verifying terms
against a signature and a repository, and updating this repository. It
supports two basic operations:
\begin{itemize}
\item $\pfunction{push}{\Sigma}{\mr,M}$ checks a small part $M$ of a
  larger term against $\Sigma$ in $\mr$, synthetizes its type $A$,
  chooses a fresh metavariable $\mmeta$ for $M$ and returns
  $\mr[\mmeta\mapsto M:A]$ and $\mmeta$.
\item $\pfunction{pull}\Sigma{\mr, \mmeta}$ returns the term $M$
  associated with $\mmeta$ in $\mr$ recursively: all metavariables are
  unfolded to their definitions.
\end{itemize}

The \emph{slicer} is the component in charge of slicing a term $M$
into many terms, pushing them to the repository to enable future
sharing, and adding version markers. It supports operations:
\begin{itemize}
\item $\pfunction{commit}\Sigma{\mr, M}$ pushes $\mathsf{vcons}\ M\
  (\mathsf{acons}\ \mmeta\ \mathsf{anil})$ to $\mr$ in several
  $\function{push}{}$ operations, where $\mmeta$ is the current head.
\item $\pfunction{merge}\Sigma{\mr, M, \mmmeta}$ pushes
  $\mathsf{vcons}\ M\ (\mathsf{acons}\ \mmeta\ (\mathsf{acons}\
  \mmmeta\ \mathsf{anil}))$ to $\mr$. Note that it doesn't actually
  perform the merge, it simply commits a previously computed merge
  node with value $M$.
\end{itemize}

The \emph{reconstructor} performs the reconstruction of the derivation
(an $M$) from the initial expression (an $e$), given the
derivations for its metavariables (an $\mr$), and the expected type
($\mathsf{exp}\ \mathsf{nil}\ \mathsf{bool}$), and commits $M$.

Finally, the \emph{compressor} computes a delta $e'$ from a
metavariable-free expression $e$ by recognizing equal subterms in
$\mr$. This can be achieved by \emph{hash-consing}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrvnat}
\bibliography{../../english.bib}

% \begin{thebibliography}{}
% \softraggedright
% \bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
% P. Q. Smith, and X. Y. Jones. ...reference text...
% \end{thebibliography}

\end{document}